### 2016-08-11 doc/paper.tex
add paper template

\documentclass{article}

\usepackage{fancyhdr}
\usepackage{hyperref}

\pagestyle{fancy}

\begin{document}

\title{ScPy -- A SuperCollider extension for performing numerical computation via embedded Python}
\author{Noah Weninger \and Abram Hindle}

\maketitle
\tableofcontents

\begin{abstract}
    this is abstract
\end{abstract}

\section{Introduction}

\section{Literature Review}

\section{Implementation}

\section{Evaluation}

\section{Conclusion}

\section{Bibilography}

\end{document}

### 2016-08-13 py/api.py
basic phase vocoder

@@ -1,5 +1,9 @@
 import numpy as np
 
+def sample_rate():
+    # TODO
+    return 44100
+
 def out(dst, src):
     """
     Write the data in src to the buffer dst.
@@ -53,3 +57,90 @@ class BackBuffer:
         Indices outsize the bounds of the buffer will wrap around.
         """
         return self.items[idx % len(self.items)]
+
+class PhaseVocoder:
+    """
+    The phase vocoder is a transformation that allows for more fine grained
+    frequency analysis and manipulation than can be achieved from a basic
+    fourier transform. It requires that some state is maintained between
+    calls, so a PhaseVocoder object should be preserved globally.
+    """
+    def __init__(self, hop):
+        """
+        Create a new phase vocoder with a given hop size. This size should be
+        the same as the STFT hop size. In SuperCollider, it is an optional
+        argument to FFT() which defaults to 1/2. For optimal quality, try 1/4
+        or 1/8.
+        """
+        self.last_phase = None
+        self.sum_phase = None
+        self.shape = None
+        self.hop = hop
+
+    def forward(self, x):
+        """
+        Performs a forwards phase vocoder transform on x. x is expected to be
+        in cartesian form. An array of complex numbers is returned where the
+        real part represents amplitude and the imaginary part represents
+        frequency.
+        """
+        if self.shape != x.shape:
+            self.shape = x.shape
+            self.last_phase = np.zeros(x.shape)
+            self.sum_phase = np.zeros(x.shape)
+            self.frame_size = x.shape[-1]
+            self.freq_per_bin = sample_rate() / self.frame_size / 2.0
+            self.step_size = self.frame_size * self.hop
+            self.expect = 2.0 * np.pi * self.hop
+            self.bins = np.arange(self.frame_size)
+        polar = to_polar(x)
+        p = polar.imag - self.last_phase
+        self.last_phase = polar.imag
+        p -= self.bins * self.expect
+        qpd = (p / np.pi).astype(int)
+        qpd = qpd + (qpd >= 0) * (qpd & 1) - (qpd < 0) * (qpd & 1)
+        p -= np.pi * qpd.astype(float)
+        p = p / self.hop / (2.0 * np.pi)
+        p = self.bins * self.freq_per_bin + p * self.freq_per_bin
+        return polar.real + p*1j
+
+    def backward(self, x):
+        """
+        Performs a backwards phase vocoder transform on x. x is expected to be
[…]

### 2016-08-13 ugen/eval.cpp
Some missing error handling

@@ -12,10 +12,22 @@ void
 importUnqualified(PyObject* main, const char* name)
 {
     PyObject* module = PyImport_ImportModule(name);
+    if (!module) {
+        PyErr_Print();
+        return;
+    }
     PyObject* attrib = PyObject_Dir(module);
+    if (!attrib) {
+        PyErr_Print();
+        return;
+    }
     Py_ssize_t numAttrs = PyList_Size(attrib);
     for (Py_ssize_t i = 0; i < numAttrs; i++) {
         PyObject* val = PyList_GetItem(attrib, i);
+        if (!val) {
+            PyErr_Print();
+            return;
+        }
         PyModule_AddObject(main, PyUnicode_AsUTF8(val),
                            PyObject_GetAttr(module, val));
     }

### 2016-08-13 ugen/ugen.cpp

@@ -128,6 +128,8 @@ Py_Ctor(Py* unit)
 {
     cout << "Py_Ctor" << endl;
     new (unit) Py;
+    if (checkError(unit))
+        return;
 
     int idx = 0;
     unit->doneAction = readAtom<int>(unit, idx);

### 2016-08-13 py/api.py
Useful pvoc offset transformations

@@ -144,3 +144,15 @@ class PhaseVocoder:
         y[row, col] += x
         return y
 
+    def to_bin_offset(self, x):
+        """
+        Converts frequency into bin relative form - the center of each bin
+        becomes that bin's zero frequency.
+        """
+        return x.real + 1j*(x.imag - self.freq_per_bin*np.indices(x.shape)[0])
+
+    def from_bin_offset(self, x):
+        """
+        The inverse of to_bin_offset().
+        """
+        return x.real + 1j*(x.imag + self.freq_per_bin*np.indices(x.shape)[0])

### 2016-08-13 ugen/eval.cpp
ignore errors, and make everything global. it's just easier to work with

@@ -79,13 +79,14 @@ Evaluator::compile(const string& code)
         if (i < line.length())
             minIndent = min(minIndent, i - 1);
     }
+    out << "seterr(all='ignore')\n";
     for (size_t i = 0; i < lines.size(); i++) {
         if (lines[i].length() > minIndent)
             out << lines[i].substr(minIndent) << "\n";
         else
             out << lines[i] << "\n";
     }
-    // out << "globals().update(locals())\n";
+    out << "globals().update(locals())\n";
     // out << "import gc\ngc.collect()\n";
 
     PyObject* obj = Py_CompileStringExFlags(out.str().c_str(), "sc-anon",

### 2016-08-14 ugen/sc_objs.cpp
remove infinities, they don't sound nice

@@ -5,7 +5,7 @@ void
 sanitize(int len, float* data)
 {
     for (int i = 0; i < len; i++) {
-        if (isnan(data[i]))
+        if (isnan(data[i]) || isinf(data[i]))
             data[i] = 0;
     }
 }

### 2016-08-16 doc/paper.tex
abstract

@@
 
-\usepackage{fancyhdr}
-\usepackage{hyperref}
+\usepackage{fancyhdr} \usepackage{hyperref}
 
@@
 \maketitle
-\tableofcontents
 
 \begin{abstract}
-    this is abstract
+    SuperCollider, a language for live coding and algorithmic synthesis of audio,
+    includes features for performing spectral analysis and manipulation of audio streams using the
+    Short-Time Fourier Transform (STFT). However, available spectral operations are limited to those
+    implemented individually as Unit Generators (UGens). Many basic operations exist and can be
+    composed to produce complex effects, however this process is tedious and leaves many of the
+    possible effects impossible to create. We present ScPy, a UGen which embeds Python scripts within
+    SuperCollider to extend the space of possible effects using the optimized vector operations
+    provided by the NumPy and SciPy libraries.
 \end{abstract}
@@
 
-\section{Bibilography}
+\section{Bibliography}
 

### 2016-08-16 doc/paper.tex
make abstract more abstract

--- 2016-08-11 doc/paper.tex
@@
 \maketitle
-\tableofcontents
 
 \begin{abstract}
-    this is abstract
+    SuperCollider, a language for sound synthesis and algorithmic composition of audio,
+    supports a wide range of synthesis, effect and analysis algorithms. However, available
+    operations are limited to those implemented explicitly as Unit Generators (UGens). Since UGens
+    are written in C/C++ and loaded as plugins during the SuperCollider server boot process, it is
+    impossible to live code UGens, which limits the user to creating sound as a composition of
+    existing UGens during a performance. Many of the vector operations required for efficiently
+    creating complex audio effects are notably missing or tedious to use. To overcome this, we
+    present ScPy, a UGen which embeds Python within SuperCollider to enable the use of the
+    optimized vector operations provided by the NumPy and SciPy libraries.
 \end{abstract}

### 2016-08-21 doc/paper.tex
write more stuff

@@
+% vim: tw=99
+
 \documentclass{article}
@@
 
-\title{ScPy -- A SuperCollider extension for performing numerical computation via embedded Python}
+\title{ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python}
+
 \author{Noah Weninger \and Abram Hindle}
@@
 \begin{abstract}
-    SuperCollider, a language for sound synthesis and algorithmic composition of audio,
-    supports a wide range of synthesis, effect and analysis algorithms. However, available
-    operations are limited to those implemented explicitly as Unit Generators (UGens). Since UGens
-    are written in C/C++ and loaded as plugins during the SuperCollider server boot process, it is
-    impossible to live code UGens, which limits the user to creating sound as a composition of
-    existing UGens during a performance. Many of the vector operations required for efficiently
-    creating complex audio effects are notably missing or tedious to use. To overcome this, we
-    present ScPy, a UGen which embeds Python within SuperCollider to enable the use of the
-    optimized vector operations provided by the NumPy and SciPy libraries.
+
+SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a wide
+range of synthesis, effect and analysis algorithms. However, available operations are limited to
+those implemented explicitly as Unit Generators (UGens). Since UGens are written in C/C++ and
+loaded as plugins during the SuperCollider server boot process, it is impossible to live code
+UGens, which limits the user to creating sound as a composition of existing UGens during a
+performance.  Many of the vector operations required for efficiently creating complex audio effects
+are notably missing or tedious to use. To overcome this, we present ScPy, a UGen which embeds
+Python within SuperCollider to enable the use of the optimized vector operations provided by the
+NumPy library.
+
 \end{abstract}
@@
 \section{Introduction}
+
+Although a number of open-source projects which interface other languages with SuperCollider exist,
+they are primarily SuperCollider clients, which have features to control synths and send other
+automation messages.  One notable exception is OctaveSC, which embeds GNU Octave within
+SuperCollider.  However, it is designed primarily for performing operations on control rate data
+arrays and its author notes performance issues which prevent it from being used on audio rate data.
+
+One of the goals of this project is to enable more flexible experimentation with FFT and phase
+vocoder based operations.  SuperCollider contains a total of 32 built in phase vocoder operations,
+which provide the ability to produce many common effects. There are additionally some extra
+operations available as user created extensions. However, many conceivable effects are impossible.
+
+ScPy allows users to overcome these limitations. By embedding the Python programming language
+within SuperCollider, it is possible to do numerical processing operations which would be too slow
+to perform in real time with SuperCollider alone. Through access to the NumPy library, previously
+difficult or impossible phase vocoder operations become trivial.  Although our usage of this
+library for testing purposes has mostly focused on FFT based operations, support exists for passing
+arbitrary data buffers to Python, which can be used to process numerical data for many other
+purposes as well.
+
+It may seem as though embedding an interpreted language within an interpreted language could offer
+no performance advantage. There is certainly some performance hit when switching languages and
+transferring data. However, the biggest advantage comes not strictly from performance but from the
+diversity of operations which become available with highly optimized implementations. NumPy is used
+internally by ScPy for handling data arrays, so the entire NumPy library of mathematical functions
+comes at no cost.  Any other Python library can easily be imported as well. Limited experiments
+have been done with the SciPy library. It would also be possible to use Theano to do processing on
+the GPU for extra performance. Implementing an equivalent to these popular Python libraries in pure
+SuperCollider is certainly possible, but it would require a very large amount of work and would see
+little widespread adoption since SuperCollider is ultimately a niche language.
 
@@
 
+Use of ScPy is through two UGens: \texttt{Py} and \texttt{PyOnce}. Both have one required argument,
+a block of Python code. A map of variable names may optionally be provided to bind SuperCollider
+variables to Python variables.  Additionally, there is a \texttt{DoneAction} argument which
+specifies an action to occur in SuperCollider after the Python code has finished executing.
+
+Since UGens may only be used as part of a synth, \texttt{PyOnce} is provided as a wrapper which
+hides this detail from the user and allows the execution of Python code outside of a synth. It is
+useful for doing initialization work.  Since the current implementation executes Python code in a
+single global namespace, a typical usage patten is for state to be defined in a \texttt{PyOnce}
+block which is later accessed and modified in a \texttt{Py} block.
+
 \section{Evaluation}
+
+\section{Future Work}
+
+At the time of writing, ScPy only supports handling audio rate data via SuperCollider's
+\texttt{Buffer} objects. In order to make some use cases more ergonomic, ScPy could certainly be
+extended to support raw audio rate input. For example, this would make time domain audio data even
+simpler to work with than frequency domain data is currently. It would also enable
+SuperCollider's FFT UGen to be easily re-implemented in Python for greater flexibility.
 

### 2016-08-23 doc/paper.tex
example in paper
typo
update paper

@@
 
-\usepackage{fancyhdr} \usepackage{hyperref}
+\usepackage{fancyhdr} \usepackage{hyperref} \usepackage{listings}
 
@@
 
-SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a wide
-range of synthesis, effect and analysis algorithms. However, available operations are limited to
-those implemented explicitly as Unit Generators (UGens). Since UGens are written in C/C++ and
-loaded as plugins during the SuperCollider server boot process, it is impossible to live code
-UGens, which limits the user to creating sound as a composition of existing UGens during a
-performance.  Many of the vector operations required for efficiently creating complex audio effects
-are notably missing or tedious to use. To overcome this, we present ScPy, a UGen which embeds
-Python within SuperCollider to enable the use of the optimized vector operations provided by the
-NumPy library.
+    SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a
+    wide range of synthesis, effect and analysis algorithms. However, available operations are
+    limited to those implemented explicitly as Unit Generators (UGens). Since UGens are written in
+    C/C++ and loaded as plugins during the SuperCollider server boot process, it is impossible to
+    live code UGens, which limits the user to creating sound as a composition of existing UGens
+    during a performance.  Many of the vector operations required for efficiently creating complex
+    audio effects are notably missing or tedious to use. To overcome this, we present ScPy, a UGen
+    which embeds Python within SuperCollider to enable the use of the optimized vector operations
+    provided by the NumPy library.
 
@@
 
-ScPy allows users to overcome these limitations. By embedding the Python programming language
+ScPy enables users to overcome these limitations. By embedding the Python programming language
 within SuperCollider, it is possible to do numerical processing operations which would be too slow
 to perform in real time with SuperCollider alone. Through access to the NumPy library, previously
-difficult or impossible phase vocoder operations become trivial.  Although our usage of this
-library for testing purposes has mostly focused on FFT based operations, support exists for passing
+difficult or impossible phase vocoder operations become trivial. Although our usage of this library
+for testing purposes has mostly focused on FFT based operations, support exists for passing
 arbitrary data buffers to Python, which can be used to process numerical data for many other
 purposes as well.
+
+\section{Methodology}
+
+During the initial planning stage of the project, we wanted to essentially provide a set of
+fundamental operations on spectral data which could be composed together to perform any conceivable
+effect. It was also a requirement that these operations would be syntactically concise in order to
+enable easy and fast experimentation. Ideally, the system would be able to handle complex effects
+in real-time.
+
+Although designing a domain-specific language (DSL) for this task was briefly considered, we
+eventually decided to use Python for a number of reasons. First, Python includes an excellent C API
+which makes embedding it within other languages trivial in comparison to many other options.
+Second, Python's syntax is easily readable and concise -- it clearly adheres to our goals. Finally,
+Python has a massive number of community built libraries for performant scientific and numerical
+computing. With a few small exceptions, these libraries contained every one of the fundamental
+spectral operations we could think of.
 
@@
 little widespread adoption since SuperCollider is ultimately a niche language.
-
-\section{Literature Review}
 
@@
 
+\lstset{ caption={SuperCollider boilerplate for no-op FFT effect with ScPy.} }
+\lstinputlisting[ basicstyle=\footnotesize\ttfamily]{../examples/template.sc}
+
+ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
+back-end, and a small Python library that provides some useful operations. Due to differences
+between these languages, connecting them was somewhat awkward in some cases.
+
+One issue involves the process of passing data between SuperCollider and C++. Since UGens are
+designed to work primarily with data streams, adding support for passing in many types of data came
+with a number of challenges. Essentially, all data must be serialized into an array of floats.
+Strings must be converted to an array of ASCII character codes, and prepended with their length.
+Arguments passed through to the Python code can have one of many types, and therefore type
+information must be encoded as well. However, just passing the type of a variable is not enough to
+know what we can do with it, since we also want to know if it inherits from a class we can use.
+
 \section{Evaluation}
@@
 extended to support raw audio rate input. For example, this would make time domain audio data even
-simpler to work with than frequency domain data is currently. It would also enable
-SuperCollider's FFT UGen to be easily re-implemented in Python for greater flexibility.
+simpler to work with than frequency domain data is currently. It would also enable SuperCollider's
+FFT UGen to be easily re-implemented in Python for greater flexibility.
 

### 2016-08-23 doc/paper.tex
use minted for code in the paper and other small improvements

@@
 
-\usepackage{fancyhdr} \usepackage{hyperref} \usepackage{listings}
+\usepackage{fancyhdr} \usepackage{hyperref} \usepackage{listings} \usepackage{minted}
 
@@
 
-\section{Implementation}
+\section{Usage}
 
@@
 a block of Python code. A map of variable names may optionally be provided to bind SuperCollider
-variables to Python variables.  Additionally, there is a \texttt{DoneAction} argument which
+variables to Python variables.  Additionally, there is an optional \texttt{DoneAction} argument which
 specifies an action to occur in SuperCollider after the Python code has finished executing.
+
+\begin{listing}[H] \inputminted[linenos=true]{SuperCollider}{../examples/template.sc}
+\caption{SuperCollider boilerplate for no-op FFT effect with ScPy.} \label{lst:boilerplate}
+\end{listing}
+
+When listing \ref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
+from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
+STFT, then finally output the resulting audio. Lines 8-9 contain Python code defining the
+processing function where operations are to be inserted. Further examples will replace only those
+lines.
+
+\section{Implementation}
 
@@
 block which is later accessed and modified in a \texttt{Py} block.
-
-\lstset{ caption={SuperCollider boilerplate for no-op FFT effect with ScPy.} }
-\lstinputlisting[ basicstyle=\footnotesize\ttfamily]{../examples/template.sc}
 
@@
 know what we can do with it, since we also want to know if it inherits from a class we can use.
-
-\section{Evaluation}
 

### 2016-08-24 doc/paper.tex
type correspondence table

@@
 
-\begin{listing}[H] \inputminted[linenos=true]{SuperCollider}{../examples/template.sc}
-\caption{SuperCollider boilerplate for no-op FFT effect with ScPy.} \label{lst:boilerplate}
+\begin{listing}[H]
+    \inputminted[linenos=true]{SuperCollider}{../examples/template.sc}
+    \caption{SuperCollider boilerplate for no-op FFT effect with ScPy.}
+    \label{lst:boilerplate}
 \end{listing}
@@
 processing function where operations are to be inserted. Further examples will replace only those
-lines.
+lines. Line 16 contains the Python code to convert the spectrum into a NumPy array, process it,
+then write back the processed version.
 
@@
 
+\begin{table}[H]
+    \caption{Type correspondence between SuperCollider and Python}
+    \begin{center}
+        \begin{tabular}{| l | l |}
+            \hline
+            SuperCollider data type & Python data type \\ \hline
+            \texttt{Float} & \texttt{float} \\ \hline
+            \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\ \hline
+            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\ \hline
+            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\ \hline
+            \texttt{Array} & \texttt{list} \\ \hline
+        \end{tabular}
+    \end{center}
+\end{table}
+
 \section{Future Work}

### 2016-08-24 ugen/ugen.cpp
return properly in weird error case

@@ -121,6 +121,7 @@ readObject(Py* unit, int& idx)
             done(unit);
             return nullptr;
     }
+    return nullptr;
 }
 
 void
@@ -189,6 +190,7 @@ bufferReady(Py* unit, int& idx)
         case Type::Unsupported:
             return false;
     }
+    return false;
 }
 
 void

### 2016-08-24 doc/paper.tex
write about the list/array conversion stuff

@@
 
+\providecommand*{\listingautorefname}{Listing}
+
 \pagestyle{fancy}
@@
 
-\title{ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python}
+\title{ScPy: A SuperCollider Extension for Perforeing Numerical Computation via Embedded Python}
 
@@
 
-When listing \ref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
+When \autoref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
 from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
@@
 
+Since \texttt{FFT} objects in SuperCollider operate on a single channel only, they must be placed
+in \texttt{Array} objects to support multichannel audio. \texttt{Array}s can hold objects of any
+type, so it is natural to convert them to Python \texttt{list} objects, which have similar
+properties. However, this presents an interesting problem: multichannel \texttt{FFT} objects are
+then converted to \texttt{lists} of \texttt{ndarray} objects in Python. These are a bit more
+awkward to work with than basic \texttt{ndarray} objects, but can easily be converted using a call
+to the \texttt{array} function in Python. This can be seen in practice on line 16 of
+\autoref{lst:boilerplate}. Although it would be possible to do this conversion
+automatically, it is left out for the purpose of not introducing a feature simply to work around an
+incomplete implementation of \texttt{FFT} objects in SuperCollider. The current conversions will
+continue to function as intended if \texttt{FFT} objects are given multichannel support in the
+future.
+
 \begin{table}[H]
@@
             \texttt{Float} & \texttt{float} \\ \hline
-            \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\ \hline
+            \texttt{UGen} (subclass of) & \texttt{ndarray} of \texttt{float} \\ \hline
             \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\ \hline

### 2016-08-24 doc/paper.tex
typos and bibilography stuff

@@
 
-\usepackage{fancyhdr} \usepackage{hyperref} \usepackage{listings} \usepackage{minted}
+\usepackage{fancyhdr}
+\usepackage{hyperref}
+\usepackage{listings}
+\usepackage{minted}
+\usepackage[backend=bibtex]{biblatex}
+
+\addbibresource{references.bib}
 
@@
 
-\title{ScPy: A SuperCollider Extension for Perforeing Numerical Computation via Embedded Python}
+\title{ScPy: A SuperCollider Extension for Performing Numerical Computation via Embedded Python}
 
@@
 
+ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
+back-end, and a small Python library that provides some useful operations. Due to differences
+between these languages, connecting them was somewhat awkward in some cases.
+
 Since UGens may only be used as part of a synth, \texttt{PyOnce} is provided as a wrapper which
@@
 block which is later accessed and modified in a \texttt{Py} block.
-
-ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
-back-end, and a small Python library that provides some useful operations. Due to differences
-between these languages, connecting them was somewhat awkward in some cases.
 
@@
 properties. However, this presents an interesting problem: multichannel \texttt{FFT} objects are
-then converted to \texttt{lists} of \texttt{ndarray} objects in Python. These are a bit more
+then converted to \texttt{list}s of \texttt{ndarray} objects in Python. These are a bit more
 awkward to work with than basic \texttt{ndarray} objects, but can easily be converted using a call
@@
 
-\section{Bibliography}
+\printbibliography
 

### 2016-08-25 doc/paper.tex
supposedly vertical rules in tables are ugly. add some references too

@@
 \usepackage{minted}
-\usepackage[backend=bibtex]{biblatex}
+\usepackage{biblatex}
+\usepackage{booktabs}
 
@@
 
-    SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a
-    wide range of synthesis, effect and analysis algorithms. However, available operations are
-    limited to those implemented explicitly as Unit Generators (UGens). Since UGens are written in
-    C/C++ and loaded as plugins during the SuperCollider server boot process, it is impossible to
-    live code UGens, which limits the user to creating sound as a composition of existing UGens
-    during a performance.  Many of the vector operations required for efficiently creating complex
-    audio effects are notably missing or tedious to use. To overcome this, we present ScPy, a UGen
-    which embeds Python within SuperCollider to enable the use of the optimized vector operations
-    provided by the NumPy library.
+    SuperCollider~\cite{wilson2011supercollider}, a language for sound synthesis and algorithmic
+    composition of audio, supports a wide range of synthesis, effect and analysis algorithms.
+    However, available operations are limited to those implemented explicitly as Unit Generators
+    (UGens). Since UGens are written in C/C++ and loaded as plugins during the SuperCollider server
+    boot process, it is impossible to live code UGens, which limits the user to creating sound as a
+    composition of existing UGens during a performance.  Many of the vector operations required for
+    efficiently creating complex audio effects are notably missing or tedious to use. To overcome
+    this, we present ScPy, a UGen which embeds Python within SuperCollider to enable the use of the
+    optimized vector operations provided by the NumPy library.
 
@@
 
-Although a number of open-source projects which interface other languages with SuperCollider exist,
-they are primarily SuperCollider clients, which have features to control synths and send other
-automation messages. One notable exception is OctaveSC, which embeds GNU Octave within
-SuperCollider. However, it is designed primarily for performing operations on control rate data
-arrays and its author notes performance issues which prevent it from being used on audio rate data.
+Although a number of open-source projects which interface other languages with SuperCollider exist
+\cite{systemsinterfacingwithsc}~\cite{magnusson2011ixi}~\cite{orlarey2009faust}, they are mostly
+SuperCollider clients, which have features to control synths and send other automation messages.
+One notable exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.
+However, it is designed primarily for performing operations on control rate data arrays and its
+author notes performance issues which prevent it from being used on audio rate data.
 
@@
 
-Although designing a domain-specific language (DSL) for this task was briefly considered, we
-eventually decided to use Python for a number of reasons. First, Python includes an excellent C API
-which makes embedding it within other languages trivial in comparison to many other options.
-Second, Python's syntax is easily readable and concise -- it clearly adheres to our goals. Finally,
-Python has a massive number of community built libraries for performant scientific and numerical
-computing. With a few small exceptions, these libraries contained every one of the fundamental
-spectral operations we could think of.
+Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
+considered, we eventually decided to use Python for a number of reasons. First, Python includes an
+excellent C API which makes embedding it within other languages trivial in comparison to many other
+options.  Second, Python's syntax is easily readable and concise --- it clearly adheres to our
+goals. Finally, Python has a massive number of community built libraries for performant scientific
+and numerical computing. With a few small exceptions, these libraries contained every one of the
+fundamental spectral operations we could think of.
 
@@
 from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
-STFT, then finally output the resulting audio. Lines 8-9 contain Python code defining the
+STFT, then finally output the resulting audio. Lines 8--9 contain Python code defining the
 processing function where operations are to be inserted. Further examples will replace only those
@@
     \begin{center}
-        \begin{tabular}{| l | l |}
-            \hline
-            SuperCollider data type & Python data type \\ \hline
-            \texttt{Float} & \texttt{float} \\ \hline
-            \texttt{UGen} (subclass of) & \texttt{ndarray} of \texttt{float} \\ \hline
-            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\ \hline
-            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\ \hline
-            \texttt{Array} & \texttt{list} \\ \hline
+        \begin{tabular}{cc}
+            \toprule
+            SuperCollider type & Python type \\
+            \midrule
+            \texttt{Float} & \texttt{float} \\
+            (subclass of) \texttt{UGen} & \texttt{ndarray} of \texttt{float} \\
+            \texttt{Buffer} & \texttt{ndarray} of \texttt{float} \\
+            \texttt{FFT} & \texttt{ndarray} of \texttt{complex} \\
+            \texttt{Array} & \texttt{list} \\
+            \bottomrule
         \end{tabular}
@@
 
-\printbibliography
+\printbibliography{}
 

### 2016-08-25 doc/paper.tex
little corrections and fixes

@@
 \usepackage{minted}
-\usepackage{biblatex}
+\usepackage[style=numeric-comp]{biblatex}
 \usepackage{booktabs}
@@
 \providecommand*{\listingautorefname}{Listing}
+\renewcommand{\sectionautorefname}{Section}
 
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
 Although a number of open-source projects which interface other languages with SuperCollider exist
-\cite{systemsinterfacingwithsc}~\cite{magnusson2011ixi}~\cite{orlarey2009faust}, they are mostly
-SuperCollider clients, which have features to control synths and send other automation messages.
-One notable exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.
-However, it is designed primarily for performing operations on control rate data arrays and its
-author notes performance issues which prevent it from being used on audio rate data.
+\cite{systemsinterfacingwithsc,magnusson2011ixi,orlarey2009faust}, they are mostly SuperCollider
+clients, which have features to control synths and send other automation messages.  One notable
+exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider.  However, it
+is designed for performing operations on control rate data arrays and suffers from performance
+issues which prevent it from being used on audio rate data.
 
@@
 vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
-which provide the ability to produce many common effects. There are additionally some extra
+which provide the ability to produce many simple effects. There are additionally some extra
 operations available as user created extensions. However, many conceivable effects are impossible.
+Some examples of novel effects can be found in~\autoref{sec:usage}.
 
@@
 
-\section{Methodology}
+\section{Methodology}\label{sec:methodology}
 
@@
 fundamental operations on spectral data which could be composed together to perform any conceivable
-effect. It was also a requirement that these operations would be syntactically concise in order to
-enable easy and fast experimentation. Ideally, the system would be able to handle complex effects
-in real-time.
+effect. It was also a requirement that these operations would be syntactically concise to enable
+easy and fast experimentation. Ideally, the system would be able to handle complex effects in
+real-time.
 
@@
 goals. Finally, Python has a massive number of community built libraries for performant scientific
-and numerical computing. With a few small exceptions, these libraries contained every one of the
-fundamental spectral operations we could think of.
+and numerical computing. With a few small exceptions, which are clearly specific to spectral data
+such as the phase vocoder, these libraries contained every one of the fundamental operations we
+could think of.
 
@@
 
-\section{Usage}
+\section{Usage}\label{sec:usage}
 
@@
 
-\section{Implementation}
+\section{Implementation}\label{sec:implementation}
 
@@
     \begin{center}
-        \begin{tabular}{cc}
+        \begin{tabular}{ll}
             \toprule

### 2016-08-26 doc/paper.tex
shepard tone example
paper improvements

@@
 \usepackage{booktabs}
+\usepackage{graphicx}
 
 \addbibresource{references.bib}
+
+\graphicspath{{images/}}
 
@@
 which provide the ability to produce many simple effects. There are additionally some extra
-operations available as user created extensions. However, many conceivable effects are impossible.
-Some examples of novel effects can be found in~\autoref{sec:usage}.
+operations available as user created extensions. However, many conceivable effects are impossible,
+including basic phase vocoder effects such as pitch shifting. Some examples of novel effects can
+be found in~\autoref{sec:usage}.
 
@@
 real-time.
+
+ScPy is unlike many New Instruments for Musical Expression (NIMEs) because it does not have the
+properties of any traditional instrument. However, it can still be considered a NIME because it
+ultimately is a tool to enhance musical expression, just with an unconventional method of
+interaction.  Dimension space analysis, as explored by Birnbaum et al.~\cite{birnbaum2005towards},
+can facilitate the comparison of use cases for NIMEs.
+
+\begin{figure}[h]
+\caption{NIME dimension space analysis for ScPy.}
+\vspace{1em}
+\includegraphics[scale=0.2]{nime-plot}
+\label{fig:nime}
+\end{figure}
 
@@
 then write back the processed version.
+
+\begin{listing}[H]
+    \begin{minted}[linenos=true]{python}
+pv = PhaseVocoder(hop)
+
+def fn(x):
+    x = pv.forward(x)
+    idx = indices(x.shape)[1]
+    x = pv.shift(x, lambda freq:
+        freq * (0.8 + mod(-time + 0.1*idx, 10)*0.045))
+    x = pv.backward(x)
+    return x
+    \end{minted}
+    \caption{A novel effect to transform any sound into a falling Shepard tone. This example makes
+    use of the phase vocoder \texttt{shift} operation, which applies a function to frequency, then
+    reorganizes the spectrum to move the new frequencies into the appropraite bins.}
+    \label{lst:shepard}
+\end{listing}
+
+\begin{listing}[H]
+    \begin{minted}[linenos=true]{python}
+pv = PhaseVocoder(hop)
+
+def fn(x):
+    x = pv.forward(x)
+    x = pv.to_bin_offset(x)
+    x.imag = -x.imag
+    x = pv.from_bin_offset(x)
+    x = pv.backward(x)
+    return x
+    \end{minted}
+    \caption{An effect which inverts the frequency of each analysis bin across the center of
+    the bin, creating an unusual detuning effect.}
+    \label{lst:binflipper}
+\end{listing}
 
@@
 
+As shown in \autoref{lst:shepard} and \autoref{lst:binflipper}, ScPy includes a phase vocoder
+implemented in Python for performing frequency transformations. Interestingly, although all of
+SuperCollider's FFT operations are referred to as phase vocoder operation, there is actually no
+implementation of a phase vocoder to be found in SuperCollider at the time of writing. It is likely
+that a phase vocoder was once intended to be implemented, but the 32 so-called phase vocoder
+effects are simply transformations to the cartesian or polar representations of the spectrum.
+
 \section{Future Work}
@@
 
+Support for more SuperCollider types, such as events or strings could additionally be added.
+
+Creating a more generic extension which supports many languages would have a number of advantages.
+One option for implementing this would be to define a specification to communicate and share data
+with a generic UGen in a running SuperCollider instance.
+
 \section{Conclusion}

### 2016-08-26 py/api.py
fix bugs in phase vocoder bin offset conversion
it might still be a bit broken, I think the frequencies are all off by 1 bin (because of DC?)

@@ -149,10 +149,10 @@ class PhaseVocoder:
         Converts frequency into bin relative form - the center of each bin
         becomes that bin's zero frequency.
         """
-        return x.real + 1j*(x.imag - self.freq_per_bin*np.indices(x.shape)[0])
+        return x.real + 1j*(x.imag - self.freq_per_bin*np.indices(x.shape)[1] + self.freq_per_bin/2)
 
     def from_bin_offset(self, x):
         """
         The inverse of to_bin_offset().
         """
-        return x.real + 1j*(x.imag + self.freq_per_bin*np.indices(x.shape)[0])
+        return x.real + 1j*(x.imag + self.freq_per_bin*np.indices(x.shape)[1] - self.freq_per_bin/2)

### 2016-08-27 doc/paper.tex
writing & editing more

@@
 \pagestyle{fancy}
+\lhead{\leftmark}
+\chead{}
+\rhead{}
 
@@
 
+ScPy is developed and maintained as an open-source project under the GPLv3 license. The
+latest code and documentation is available on GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+
 \section{Methodology}\label{sec:methodology}
@@
 
-Use of ScPy is through two UGens: \texttt{Py} and \texttt{PyOnce}. Both have one required argument,
-a block of Python code. A map of variable names may optionally be provided to bind SuperCollider
-variables to Python variables.  Additionally, there is an optional \texttt{DoneAction} argument which
-specifies an action to occur in SuperCollider after the Python code has finished executing.
+Use of ScPy is through the classes \texttt{Py} and \texttt{PyOnce}. \texttt{Py} is used for
+processing real-time data, and \texttt{PyOnce} is used for doing initialization or other one time
+processing. Both have one required argument, a block of Python code. A map of variable names may
+optionally be provided to bind SuperCollider variables to Python variables.  Additionally, there is
+an optional \texttt{DoneAction} argument which specifies an action to occur in SuperCollider after
+the Python code has finished executing.
 
@@
     use of the phase vocoder \texttt{shift} operation, which applies a function to frequency, then
-    reorganizes the spectrum to move the new frequencies into the appropraite bins.}
+    reorganizes the spectrum to move the new frequencies into the appropriate bins.}
     \label{lst:shepard}
@@
 implemented in Python for performing frequency transformations. Interestingly, although all of
-SuperCollider's FFT operations are referred to as phase vocoder operation, there is actually no
-implementation of a phase vocoder to be found in SuperCollider at the time of writing. It is likely
-that a phase vocoder was once intended to be implemented, but the 32 so-called phase vocoder
-effects are simply transformations to the cartesian or polar representations of the spectrum.
+SuperCollider's FFT operations are referred to as phase vocoder operations, there is actually no
+implementation of a phase vocoder to be found in the SuperCollider code base at the time of
+writing. It is likely that a phase vocoder was once intended to be implemented, but the 32
+so-called phase vocoder effects are simply transformations to the Cartesian or polar
+representations of the spectrum.
 
@@
 
-Support for more SuperCollider types, such as events or strings could additionally be added.
+Support for more SuperCollider types, such as events or strings, could be added to open up ScPy to
+more use cases.
+
+Multi-core processing support would allow for many more sounds to be processed in real-time.
 
@@
 
-\section{Conclusion}
+Although extending SuperCollider was the focus of this project, similar improvements could be made
+to the CSound\cite{boulanger2000csound,lazzarini2005extensions}, Chuck\cite{wang2008chuck} or
+Faust\cite{orlarey2009faust} environments as well.
 

### 2016-08-27 doc/paper.tex
hello world and minor corrections

@@
 and numerical computing. With a few small exceptions, which are clearly specific to spectral data
-such as the phase vocoder, these libraries contained every one of the fundamental operations we
+(such as the phase vocoder), these libraries contained every one of the fundamental operations we
 could think of.
@@
 SuperCollider is certainly possible, but it would require a very large amount of work and would see
-little widespread adoption since SuperCollider is ultimately a niche language.
+no adoption outside of the audio computing community.
 
@@
 the Python code has finished executing.
+
+\begin{listing}[H]
+    \begin{minted}[linenos=true]{SuperCollider}
+(PyOnce("
+    # comments in Python have different syntax!
+    print('Hello from Python!')
+    print('x is', x)
+", (x: 62)))
+    \end{minted}
+    \caption{Hello world with ScPy.}
+    \label{lst:helloworld}
+\end{listing}
 

### 2016-08-27 doc/paper.tex
out function doc
little fixes

@@
 
-ScPy is developed and maintained as an open-source project under the GPLv3 license. The
-latest code and documentation is available on GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest code
+and documentation is available on GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
 
@@
 When \autoref{lst:boilerplate} is run, an anonymous synth is created which will take in audio
-from an external source, perform a STFT, process the spectral data with ScPy, perform an inverse
+from an external source, perform an STFT, process the spectral data with ScPy, perform an inverse
 STFT, then finally output the resulting audio. Lines 8--9 contain Python code defining the
@@
 then write back the processed version.
+
+A special function, \texttt{out}, is provided to mutate data in SuperCollider buffers. It takes a
+variable which is bound to a \texttt{Buffer} or \texttt{FFT} object and a new value. The buffer is
+modified in place to hold the new data.
 

### 2016-08-28 doc/paper.tex
some lit review + fixes

@@
 
-    SuperCollider~\cite{wilson2011supercollider}, a language for sound synthesis and algorithmic
-    composition of audio, supports a wide range of synthesis, effect and analysis algorithms.
-    However, available operations are limited to those implemented explicitly as Unit Generators
-    (UGens). Since UGens are written in C/C++ and loaded as plugins during the SuperCollider server
-    boot process, it is impossible to live code UGens, which limits the user to creating sound as a
-    composition of existing UGens during a performance.  Many of the vector operations required for
-    efficiently creating complex audio effects are notably missing or tedious to use. To overcome
-    this, we present ScPy, a UGen which embeds Python within SuperCollider to enable the use of the
-    optimized vector operations provided by the NumPy library.
+    SuperCollider, a language for sound synthesis and algorithmic composition of audio, supports a
+    wide range of synthesis, effect and analysis algorithms.  However, available operations are
+    limited to those implemented explicitly as Unit Generators (UGens). Since UGens are written in
+    C/C++ and loaded as plugins during the SuperCollider server boot process, it is impossible to
+    live code UGens, which limits the user to creating sound as a composition of existing UGens
+    during a performance.  Many of the vector operations required for efficiently creating complex
+    audio effects are notably missing or tedious to use. To overcome this, we present ScPy, a UGen
+    which embeds Python within SuperCollider to enable the use of the optimized vector operations
+    provided by the NumPy library.
 
@@
 
+\section{Literature Review}\label{sec:lit}
+
+With any New Interface for Musical Expression (NIME), it is important to evaluate how it performs
+and is interacted with across various dimensions. Birnbaum et al.~\cite{birnbaum2005towards}
+present a new model for evaluating musical instruments by determining the position of the
+instrument along 7 axes: \textit{Required Expertise}, \textit{Musical Control}, \textit{Feedback
+Modalities}, \textit{Degrees of Freedom}, \textit{Inter-actors}, \textit{Distribution in Space},
+and the \textit{Role of Sound}. Some of these axes represent a continuous spectrum, where as others
+have discrete steps. These dimensions are arranged to form a heptagon in the 2D plane, where each
+corner of the shape is placed at a distance to the center proportional to the score of the
+instrument being evaluated along the corresponding axis. The specific layout of the dimensions used
+in the paper additionally has properties which allow instruments of different classes to be easily
+distinguished. Plots of instruments tend to occupy the right side of the graph, where as
+installations tend to occupy the left. Although this model performs well on the instruments
+evaluated in the paper, it is based on subjective evaluation and it is likely that it does not
+adequately cover the space of all possible instruments.
+
+ScPy is unlike many NIMEs because it does not have the properties of any traditional instrument.
+However, it can still be considered a NIME because it ultimately is a tool for musical expression,
+just with an unconventional method of interaction. A plot of ScPy in the 7-dimensional space
+developed by Birnbaum et al.\ can be seen in~\autoref{fig:nime}.
+
+\begin{figure}[ht]
+\caption{NIME dimension space analysis for ScPy.}
+\vspace{1em}
+\includegraphics[scale=0.2]{nime-plot}
+\label{fig:nime}
+\end{figure}
+
+The phase vocoder is an essential part of spectral audio manipulation. De G{\"o}tzen et
+al.~\cite{de2000traditional} present an open source Matlab implementation along with a number of
+tricks to make the implementation faster or simpler. The implementation is split into two
+functions, \texttt{pv\_analyze} and \texttt{pv\_synthesize}. These functions operate as a tightly
+coupled set due to the phase information maintained between frames. One of the techniques presented
+is to perform a \texttt{fftshift} operation in order to simplify the phase relationship between
+neighbouring bins. Another is to zero pad frames: by adding extra zeros on each side of a frame, a
+larger FFT size can be used, producing better accuracy while maintaining good time resolution.
+
+The code in this paper was used as a reference for the phase vocoder implementation in ScPy.
+Additionally, some of the extra techniques mentioned could improve the quality or flexibility of
+the implementation in the future.
+
+Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
+code to generate sound is written live during a performance, often on a laptop computer. Although
+the use graphical programming environments such as Reaktor, Pure Data, or MAX/MSP in live
+performances is in some sense live coding, the paper focuses on the use of text based programming
+environments. Many of the pros and cons of live coding are presented. Although live coding adds
+some very interesting new possibilities to live music performance, it is not without it's
+challenges. Case studies of two live coding environments are included: Slub and SuperCollider with
+the just in time library.
+
+ScPy can be used for live coding of audio effects, similarly to how the environments studied in
+this paper can be used to live code synths and pattens. However, at the time of writing one issue
+is blocking ScPy from realistically being used for live coding performances: the Python global
+environment can be put into a state which prevents the code from executing without throwing an
+exception, which could halt a performance in its tracks if the programmer made one of a number of
+errors. This may be avoidable if more advanced error detection is implemented in the future.
+
+Live coding presents an interesting contrast to contemporary music with many clear differences.
+Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
+practice. While some argue that live coding must be performed live in front of an
+audience, an alternate perspective suggests that time constraints are not necessary and that the
+practice can refer to any improvisational music created by programming. The practice of live coding
+is still quite young --- it is mentioned that professional violinists will typically have practiced
+for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
+hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
+extensive programming experience. Another thing mentioned is the high level of abstraction offered
+by live coding creates a large startup time for producing meaningful output from a blank slate.
+Additionally, many exercises are presented for improving one's live coding ability.
+
+The ideas in Nilson's paper shed light on some of the challenges faced in this project.
+One such challenge is the amount of writing required to produce an interesting result from a blank
+slate. We intend to minimise this by providing many effect primitives and composition
+mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
+quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
+
+While the Fourier transform is by far the most commonly used technique for spectral manipulation in
+digital audio effects, there exist numerous other transformations which may be useful as well to
+achieve certain audio effects. One such technique is the Mellin transform, described in a audio
+effects context by De Sena and Rocchesso~\cite{de2004fast}. The Mellin transform has a scale
+invariance property that is analogous to the shift invariance property of the Fourier transform. It
+can be computed quickly by applying two extra steps before doing a Fourier transform: exponential
+time warping and exponential multiplication. A few simple audio effects are presented using the
+Mellin transform, time stretching, filtering and random phase deviation (whisperization).
+
+As the Mellin transform can be implemented simply as an extension to the Fourier transform, it
+should be fairly easy to add support for it in ScPy, and it will hopefully enable the use of
+some different audio effects than can be achieved easily with the Fourier transform. However, the
+paper mentions that a short-time Mellin transform has not been attempted, so some further work may
+need to be done in that area first.
+
+Domain-Specific Languages (DSLs) are small programming languages designed for use in a specific
+problem domain. They often lack general purpose programming constructs and have a syntax which is
+very concise for describing programs in their domain. Van Deursen et al.~\cite{van2000domain}
+present a survey of available DSL literature, along with the pros and cons of using a DSL, example
+DSLs, design methodology and implementation approaches. The biggest advantage of DSLs is that they
+enable expression of solutions in a way that is common for the problem domain. However, they can be
+costly to develop and difficult to design. Implementation techniques include the classical approach
+of interpretation or compilation, embedded languages / domain-specific libraries, preprocessing, or
+compiler extension. Additionally, the references of the paper each include a short summary.
+
+This review reveals many of the possible options for the DSL used within the project. Although we
+decided not to design a DSL for our project, analyzing the design and implementation techniques
+presented here will be helpful for future work in this area.
+
+Klingbeil~\cite{klingbeil2005software} describes the techniques used in a high quality graphical
+spectral editing and resynthesis program called SPEAR\@. The FFT analysis is done taking into
+account the fundamental frequency of the sound to analyze, so that the transform can be optimized
+to accurately capture as many partials as possible. Spectral peaks are then selected and very quiet
+peaks are discarded. A linear prediction method is used to follow the peaks over time. The data
+structure for storage of spectral data involves division of time into a number of frames which keep
+track of which partials are active during that time, so that the search for a specific partial can
+be narrowed. User interface design choices are also covered in detail.
+
+Although the interface and granularity of SPEAR differs significantly from those in project, the
+techniques for high quality resynthesis presented should prove useful for improving the quality of
+the sound output.
+
+Keiler and Sylvain~\cite{keiler2002survey} present a survey of 6 techniques for accurate analysis
+of stationary sounds. The first method, plain FFT, uses a Hann window and assumes the frequencies
+of the sinusoids at centered within each bin. Parabolic interpolation notices that the shape of a
+magnitude peak looks like a parabola, and thus attempts to fit a parabola to the magnitudes of bins
+nearby each local maxima, yielding more accurate results than a plain FFT\@. The triangle algorithm
+is similar to the parabolic interpolation method but fits a triangular function instead of a
+parabola. Spectral reassignment uses the derivative of the analysis window to estimate the true
+values of the bins as the center of gravity of the energy of each bin. The derivative algorithm
+also uses the derivative of the analysis window to determine both a more accurate frequency and
+amplitude. It can be further extended to use the first nth derivatives. The phase vocoder uses the
+phase information of two successive FFTs to improve the frequency resolution by calculating the
+phase derivative. Comparison of the techniques found that the spectral reassignment, derivative
+algorithm, and phase vocoder methods had the best performance. Surprisingly, those three techniques
+produced almost identical results, suggesting that they may have equivalent theoretical basis.
+
+The two other techniques that performed similarly to the phase vocoder will be interesting to
+analyze in more detail since they may present a solution to the \textit{phasiness} problem of the
+phase vocoder due their lack of persistent state. However, they require extra processing so it is
+not clear if the trade-off will be worthwhile.
+
+Audio morphing is defined as a gradual change from one sound to the other. While morphing can be
+achieved to some extent simply by cross fading two signals, perceptually this sounds simply like
+one sound is becoming quieter as the other is becoming louder. Sethares and Bucklew
+\cite{sethares2012kernel} propose density and kernel based techniques that aim to provide a more
+convincing morph operation where there is no clear separation between the two sounds at any point
+during the morphing process. The method allows the use of different kernel functions to provide
+different kinds of morphing. These include \textit{Cross-fade}, \textit{Heat equation}, and
+\textit{Harmonic integrity}. The morphing technique is implemented using a phase vocoder.
+
+As a future extension to ScPy, these morphing techniques could be implemented in Python.
+
+As Python is an easily extendible general purpose language, it has applications in many scientific
+fields.  Glover et al.~\cite{glover2011python} present a discussion of the use of Python for audio
+signal processing. They note that while audio processing libraries exist for many languages,
+scripting languages such as Python or MATLAB see widespread adoption because of the speed at which
+prototypes can be built. General purpose Python libraries such as NumPy, SciPy and Matplotlib
+include many essential functions for audio processing, such as file IO, Fourier transforms, signal
+processing and interpolation. There are additionally more specific libraries such as Simpl, a
+sinusoidal modelling library, and Modal, an onset detection library.
+
+Any of these libraries can easily be used within ScPy simply by installing and importing them.
+
+Another open-source Python library which could be used in ScPy is librosa~\cite{mcfee2015librosa}
+by McFee et al.\ librosa is a music signal analysis library designed to help transition music
+information retrieval (MIR) researchers into using Python, as well as make MIR techniques readily
+available to all Python programmers.\ librosa's core design choices and conventions are documented
+in detail to inspire future library authors to follow similarly good coding practice. As well as
+standard audio analysis functions, librosa also includes support for effects and audio IO.\@
+
+librosa has quite an extensive list of features and would likely cover all of the audio related
+functionality which is missing from NumPy. It can be used with ScPy simply by installing and importing
+it.
+
+The ixi lang~\cite{magnusson2011ixi} is a live coding language. It is implemented as an interpreter
+built in SuperCollider, a choice which allows for SuperCollider and ixi lang code to be mixed
+together in a similar manner to ScPy. The ixi lang is built to reduce the startup time in a live
+coding performance and enable faster musical experimentation through syntax more concise than
+SuperCollider. The paper also contains user feedback on the usability of the ixi lang.
+
+Since the ixi lang is similar to ScPy in many ways, many of the design choices made in ixi lang are
+also appropriate in ScPy. Additionally, lessons learned from the user feedback for ixi lang could
+potentially be used to improve usability of ScPy.
+
 \section{Methodology}\label{sec:methodology}
@@
 real-time.
-
-ScPy is unlike many New Instruments for Musical Expression (NIMEs) because it does not have the
-properties of any traditional instrument. However, it can still be considered a NIME because it
-ultimately is a tool to enhance musical expression, just with an unconventional method of
-interaction.  Dimension space analysis, as explored by Birnbaum et al.~\cite{birnbaum2005towards},
-can facilitate the comparison of use cases for NIMEs.
-
-\begin{figure}[h]
-\caption{NIME dimension space analysis for ScPy.}
-\vspace{1em}
-\includegraphics[scale=0.2]{nime-plot}
-\label{fig:nime}
-\end{figure}
 
@@
 
+In order to facilitate the use of ScPy in live coding performances, many errors are ignored or
+handled in a way that can be easily recovered from. In particular, floating point errors such as
+division by zero or other invalid operations are ignored. Special values such as infinity or NAN
+are replaced with zeros. This is done because SuperCollider has poor handling of these special
+values and in some cases will crash if they are found. Python exceptions are printed to the
+SuperCollider post window, where they can easily be examined and corrected by the user.
+
 \section{Usage}\label{sec:usage}
@@
 
-\begin{listing}[H]
+\begin{listing}[ht]
     \begin{minted}[linenos=true]{SuperCollider}
@@
 
-\begin{listing}[H]
+\begin{listing}[ht]
     \inputminted[linenos=true]{SuperCollider}{../examples/template.sc}
@@
 
-\begin{listing}[H]
+\begin{listing}[ht]
     \begin{minted}[linenos=true]{python}
@@
 
-\begin{table}[H]
+\begin{table}[ht]
     \caption{Type correspondence between SuperCollider and Python}
@@
 representations of the spectrum.
+
+TODO data flow diagram
 

### 2016-08-28 doc/paper.tex
editing, writing, and diagram

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+
+\usetikzlibrary{arrows}
 
@@
 One of the goals of this project is to enable more flexible experimentation with FFT and phase
-vocoder based operations. SuperCollider contains a total of 32 built in phase vocoder operations,
+vocoder based operations. SuperCollider contains a total of 32 built in spectral operations,
 which provide the ability to produce many simple effects. There are additionally some extra
@@
 different kinds of morphing. These include \textit{Cross-fade}, \textit{Heat equation}, and
-\textit{Harmonic integrity}. The morphing technique is implemented using a phase vocoder.
-
-As a future extension to ScPy, these morphing techniques could be implemented in Python.
+\textit{Harmonic integrity}. The morphing technique is implemented using a phase vocoder. These
+techniques could be implemented in Python as a future extension to ScPy.
 
@@
 processing and interpolation. There are additionally more specific libraries such as Simpl, a
-sinusoidal modelling library, and Modal, an onset detection library.
-
-Any of these libraries can easily be used within ScPy simply by installing and importing them.
+sinusoidal modelling library, and Modal, an onset detection library. Any of these libraries can
+easily be used within ScPy simply by installing and importing them.
 
@@
 
+Control UGens only provide a single floating point value per cycle so it would seem natural to
+convert them to a Python \texttt{float}, but in ScPy they are instead transferred to Python in a
+one element \texttt{ndarray}. This is because Python provides no numerical types with interior
+mutability. In practice, NumPy accepts a one element \texttt{ndarray} almost anywhere a
+\texttt{float} is accepted, so this detail is not an annoyance.
+
 \begin{table}[ht]
@@
 
+\begin{figure}[ht]
+    \caption{A high level overview of data flow and processing.}
+    \begin{center}
+        \begin{tikzpicture}[
+            every matrix/.style={ampersand replacement=\&,column sep=2cm,row sep=2cm},
+            sink/.style={draw,thick,rounded corners,fill=green!20,inner sep=.3cm},
+            process/.style={draw,thick,rounded corners,fill=blue!20,inner sep=.3cm},
+            datastore/.style={draw,thick,inner sep=.3cm},
+            to/.style={->,>=stealth',shorten >=1pt,semithick},
+            both/.style={<->,>=stealth',shorten >=1pt,semithick},
+            every node/.style={align=center}]
+
+            \matrix{\node[sink] (sc) {SuperCollider}; \\
+                    \node[datastore] (objects) {buffer}; \& \node[process] (scpy) {ScPy}; \\
+                    \& \node[sink] (py) {Python}; \\
+            };
+
+            \draw[to] (sc) -| node[near end,right] {next()} (scpy);
+            \draw[to] (objects) |- node[near start,left] {data pointer} (py);
+            \draw[both] (objects) -- node[midway,left] {copy data} (sc);
+            \draw[to] (scpy) -- node[midway,right] {execute()} (py);
+            \draw[to] (scpy) -- node[midway,above] {update()} (objects);
+        \end{tikzpicture}
+    \end{center}
+\end{figure}
+
 As shown in \autoref{lst:shepard} and \autoref{lst:binflipper}, ScPy includes a phase vocoder
@@
 representations of the spectrum.
-
-TODO data flow diagram
 
@@
 Support for more SuperCollider types, such as events or strings, could be added to open up ScPy to
-more use cases.
+more use cases. Adding support for more types should be trivial in the existing code base.
 
 Multi-core processing support would allow for many more sounds to be processed in real-time.
+Currently, all processing is done on a single thread, even with many UGens running at once. Using
+the supernova server~\cite{blechmann2005supernova} may be helpful, however it has not yet been
+tested and ScPy may require some modifications to be usable with supernova.
 
 Creating a more generic extension which supports many languages would have a number of advantages.
-One option for implementing this would be to define a specification to communicate and share data
-with a generic UGen in a running SuperCollider instance.
+Although extending SuperCollider was the focus of this project, similar improvements could be made
+to the CSound~\cite{boulanger2000csound,lazzarini2005extensions}, Chuck~\cite{wang2008chuck} or
+Faust~\cite{orlarey2009faust} environments as well. It may be desirable to completely separate the
+embedded language from the host, in order to allow easier interoperation and a cleaner
+implementation.
 
-Although extending SuperCollider was the focus of this project, similar improvements could be made
-to the CSound\cite{boulanger2000csound,lazzarini2005extensions}, Chuck\cite{wang2008chuck} or
-Faust\cite{orlarey2009faust} environments as well.
-
-\printbibliography{}
+\newpage\printbibliography{}
 

### 2016-08-28 doc/paper.tex
radial graphs in latex are easy!

@@
 \usepackage{booktabs}
-\usepackage{graphicx}
 \usepackage{tikz}
+\usepackage{tkz-kiviat}
 
@@
 \begin{figure}[ht]
-\caption{NIME dimension space analysis for ScPy.}
-\vspace{1em}
-\includegraphics[scale=0.2]{nime-plot}
-\label{fig:nime}
+    \caption{NIME dimension space analysis for ScPy.}\label{fig:nime}
+    \begin{center}
+        \begin{tikzpicture}[rotate=-116,xscale=-1]
+            \tkzKiviatDiagram[label distance=.5cm,
+                    radial=7,
+                    gap=0.75,
+                    lattice=4]{Required Expertise,
+                    Musical Control,
+                    Degrees of Freedom (Inputs),
+                    Feedback Modalities (Outputs),
+                    Inter-actors,
+                    Distribution in Space,
+                    Role of Sound}
+            \tkzKiviatLine[thick,color=blue,mark=none,
+                           fill=blue!20,opacity=.5](4,4,1.5,1.5,0.5,0.5,3)
+        \end{tikzpicture}
+    \end{center}
 \end{figure}

### 2016-08-28 doc/paper.tex
more literature

@@
             \tkzKiviatLine[thick,color=blue,mark=none,
-                           fill=blue!20,opacity=.5](4,4,1.5,1.5,0.5,0.5,3)
+                           fill=blue,opacity=.5](4,4,1.5,1.5,0.5,0.5,3)
         \end{tikzpicture}
@@
 ScPy can be used for live coding of audio effects, similarly to how the environments studied in
-this paper can be used to live code synths and pattens. However, at the time of writing one issue
-is blocking ScPy from realistically being used for live coding performances: the Python global
-environment can be put into a state which prevents the code from executing without throwing an
-exception, which could halt a performance in its tracks if the programmer made one of a number of
-errors. This may be avoidable if more advanced error detection is implemented in the future.
+this paper can be used to live code synths and pattens. More details about how ScPy was designed to
+facilitate this can be found in \autoref{sec:methodology}.
 
@@
 
+Laroche and Dolson~\cite{laroche1999improved} present an improvement to the phase vocoder which
+aims to prevent the ``phasiness'' problem where different parts of a sound lose their phase
+coherence, as well as provide a significant decrease in computational cost. The paper recognizes
+that there are two types of phase coherence that must be preserved in any STFT based algorithm to
+ensure that the reproduced sound is without artifacts: horizontal and vertical. The basic phase
+vocoder algorithm is built to preserve horizontal phase coherence --- but it does so at the expense
+of vertical phase coherence. To remedy this, they build on prior work by presenting two new
+phase-locking techniques.
+
+Similar techniques~\cite{laroche1999new} enable new phase vocoder techniques that enable
+exotic audio effects. The core of the techniques is a peak detection process followed by peak
+shifting. Since these new techniques operate on peaks, the standard phase vocoder algorithm can be
+simplified to not require phase unwrapping, reducing computational complexity. Additionally, they
+produce better quality output than the standard phase vocoder algorithm because of implicit
+phase-locking.
+
+The phase vocoder currently implemented in ScPy uses the standard algorithm. Future experiments
+could implement many of the proposed modifications to the phase vocoder to improve sound quality.
+
+Bradford et al.\ have published a number of works detailing the sliding
+DFT~\cite{bradford2005sliding} and associated sliding phase vocoder~\cite{bradford2007sliding}.
+These new algorithms are efficient implementations for the STFT with a hop size of a single sample.
+Although such a small hop size would be impossible to process in real time using current hardware
+and the STFT, the sliding DFT makes it possible using consumer GPU
+hardware~\cite{bradford2011real}. The sliding phase vocoder additionally has a number of advantges
+over traditional implementations. Quality is greatly improved (at a high computational cost). Each
+analysis bin can hold any frequency, essentially making the spectrum more of a generic oscillator
+bank than a traditional spectrum. This simplifies effects such as pitch shifting to a simple vector
+multiplication in the frequency domain. Additionally, since the time resolution is so fine, effects
+that require time modulation such as FM or AM can now be performed in the frequency domain.
+
+Using Theano, the sliding DFT and sliding phase vocoder could be implemented easily in ScPy. It
+remains unknown if it will be possible to run them in real time, but even if not this future
+extension could add the possibility to experiment with many novel effects.
+
 \section{Methodology}\label{sec:methodology}
@@
 
-Although designing a domain-specific language (DSL)~\cite{van2000domain} for this task was briefly
-considered, we eventually decided to use Python for a number of reasons. First, Python includes an
-excellent C API which makes embedding it within other languages trivial in comparison to many other
-options. Second, Python's syntax is easily readable and concise --- it clearly adheres to our
-goals. Finally, Python has a massive number of community built libraries for performant scientific
-and numerical computing. With a few small exceptions, which are clearly specific to spectral data
-(such as the phase vocoder), these libraries contained every one of the fundamental operations we
-could think of.
+Although designing a domain-specific language for this task was briefly considered, we eventually
+decided to use Python for a number of reasons. First, Python includes an excellent C API which
+makes embedding it within other languages trivial in comparison to many other options. Second,
+Python's syntax is easily readable and concise --- it clearly adheres to our goals. Finally, Python
+has a massive number of community built libraries for performant scientific and numerical
+computing. With a few small exceptions, which are clearly specific to spectral data (such as the
+phase vocoder), these libraries contained every one of the fundamental operations we could think
+of.
 
@@
 handled in a way that can be easily recovered from. In particular, floating point errors such as
-division by zero or other invalid operations are ignored. Special values such as infinity or NAN
+division by zero or other invalid operations are ignored. Special values such as infinity or NaN
 are replaced with zeros. This is done because SuperCollider has poor handling of these special
@@
 
+\begin{figure}[ht]
+    \caption{A high level overview of data flow and processing.}
+    \begin{center}
+        \begin{tikzpicture}[
+            every matrix/.style={ampersand replacement=\&,column sep=2cm,row sep=2cm},
+            sink/.style={draw,thick,rounded corners,fill=green!20,inner sep=.3cm},
+            process/.style={draw,thick,rounded corners,fill=blue!20,inner sep=.3cm},
+            datastore/.style={draw,thick,inner sep=.3cm},
+            to/.style={->,>=stealth',shorten >=1pt,semithick},
+            both/.style={<->,>=stealth',shorten >=1pt,semithick},
+            every node/.style={align=center}]
+
+            \matrix{\node[sink] (sc) {SuperCollider}; \\
+                    \node[datastore] (objects) {buffer}; \& \node[process] (scpy) {ScPy}; \\
+                    \& \node[sink] (py) {Python}; \\
+            };
+
+            \draw[to] (sc) -| node[near end,right] {next()} (scpy);
+            \draw[to] (objects) |- node[near start,left] {data pointer} (py);
+            \draw[both] (objects) -- node[midway,left] {copy data} (sc);
+            \draw[to] (scpy) -- node[midway,right] {execute()} (py);
+            \draw[to] (scpy) -- node[midway,above] {update()} (objects);
+        \end{tikzpicture}
+    \end{center}
+\end{figure}
+
 ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
@@
 
-\begin{figure}[ht]
-    \caption{A high level overview of data flow and processing.}
-    \begin{center}
-        \begin{tikzpicture}[
-            every matrix/.style={ampersand replacement=\&,column sep=2cm,row sep=2cm},
-            sink/.style={draw,thick,rounded corners,fill=green!20,inner sep=.3cm},
-            process/.style={draw,thick,rounded corners,fill=blue!20,inner sep=.3cm},
-            datastore/.style={draw,thick,inner sep=.3cm},
-            to/.style={->,>=stealth',shorten >=1pt,semithick},
-            both/.style={<->,>=stealth',shorten >=1pt,semithick},
-            every node/.style={align=center}]
-
-            \matrix{\node[sink] (sc) {SuperCollider}; \\
-                    \node[datastore] (objects) {buffer}; \& \node[process] (scpy) {ScPy}; \\
-                    \& \node[sink] (py) {Python}; \\
-            };
-
-            \draw[to] (sc) -| node[near end,right] {next()} (scpy);
-            \draw[to] (objects) |- node[near start,left] {data pointer} (py);
-            \draw[both] (objects) -- node[midway,left] {copy data} (sc);
-            \draw[to] (scpy) -- node[midway,right] {execute()} (py);
-            \draw[to] (scpy) -- node[midway,above] {update()} (objects);
-        \end{tikzpicture}
-    \end{center}
-\end{figure}
-
 As shown in \autoref{lst:shepard} and \autoref{lst:binflipper}, ScPy includes a phase vocoder

### 2016-08-28 doc/paper.tex
lit shuffling

@@
 exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider. However, it
-is designed for performing operations on control rate data arrays and suffers from performance
+is designed for performing operations on control rate data matrices and suffers from performance
 issues which prevent it from being used on audio rate data.
@@
 
-Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
-code to generate sound is written live during a performance, often on a laptop computer. Although
-the use graphical programming environments such as Reaktor, Pure Data, or MAX/MSP in live
-performances is in some sense live coding, the paper focuses on the use of text based programming
-environments. Many of the pros and cons of live coding are presented. Although live coding adds
-some very interesting new possibilities to live music performance, it is not without it's
-challenges. Case studies of two live coding environments are included: Slub and SuperCollider with
-the just in time library.
-
-ScPy can be used for live coding of audio effects, similarly to how the environments studied in
-this paper can be used to live code synths and pattens. More details about how ScPy was designed to
-facilitate this can be found in \autoref{sec:methodology}.
-
-Live coding presents an interesting contrast to contemporary music with many clear differences.
-Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
-practice. While some argue that live coding must be performed live in front of an
-audience, an alternate perspective suggests that time constraints are not necessary and that the
-practice can refer to any improvisational music created by programming. The practice of live coding
-is still quite young --- it is mentioned that professional violinists will typically have practiced
-for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
-hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
-extensive programming experience. Another thing mentioned is the high level of abstraction offered
-by live coding creates a large startup time for producing meaningful output from a blank slate.
-Additionally, many exercises are presented for improving one's live coding ability.
-
-The ideas in Nilson's paper shed light on some of the challenges faced in this project.
-One such challenge is the amount of writing required to produce an interesting result from a blank
-slate. We intend to minimise this by providing many effect primitives and composition
-mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
-quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
-
-While the Fourier transform is by far the most commonly used technique for spectral manipulation in
-digital audio effects, there exist numerous other transformations which may be useful as well to
-achieve certain audio effects. One such technique is the Mellin transform, described in a audio
-effects context by De Sena and Rocchesso~\cite{de2004fast}. The Mellin transform has a scale
-invariance property that is analogous to the shift invariance property of the Fourier transform. It
-can be computed quickly by applying two extra steps before doing a Fourier transform: exponential
-time warping and exponential multiplication. A few simple audio effects are presented using the
-Mellin transform, time stretching, filtering and random phase deviation (whisperization).
-
-As the Mellin transform can be implemented simply as an extension to the Fourier transform, it
-should be fairly easy to add support for it in ScPy, and it will hopefully enable the use of
-some different audio effects than can be achieved easily with the Fourier transform. However, the
-paper mentions that a short-time Mellin transform has not been attempted, so some further work may
-need to be done in that area first.
-
-Domain-Specific Languages (DSLs) are small programming languages designed for use in a specific
-problem domain. They often lack general purpose programming constructs and have a syntax which is
-very concise for describing programs in their domain. Van Deursen et al.~\cite{van2000domain}
-present a survey of available DSL literature, along with the pros and cons of using a DSL, example
-DSLs, design methodology and implementation approaches. The biggest advantage of DSLs is that they
-enable expression of solutions in a way that is common for the problem domain. However, they can be
-costly to develop and difficult to design. Implementation techniques include the classical approach
-of interpretation or compilation, embedded languages / domain-specific libraries, preprocessing, or
-compiler extension. Additionally, the references of the paper each include a short summary.
-
-This review reveals many of the possible options for the DSL used within the project. Although we
-decided not to design a DSL for our project, analyzing the design and implementation techniques
-presented here will be helpful for future work in this area.
-
-Klingbeil~\cite{klingbeil2005software} describes the techniques used in a high quality graphical
-spectral editing and resynthesis program called SPEAR\@. The FFT analysis is done taking into
-account the fundamental frequency of the sound to analyze, so that the transform can be optimized
-to accurately capture as many partials as possible. Spectral peaks are then selected and very quiet
-peaks are discarded. A linear prediction method is used to follow the peaks over time. The data
-structure for storage of spectral data involves division of time into a number of frames which keep
-track of which partials are active during that time, so that the search for a specific partial can
-be narrowed. User interface design choices are also covered in detail.
-
-Although the interface and granularity of SPEAR differs significantly from those in project, the
-techniques for high quality resynthesis presented should prove useful for improving the quality of
-the sound output.
-
 Keiler and Sylvain~\cite{keiler2002survey} present a survey of 6 techniques for accurate analysis
@@
 techniques could be implemented in Python as a future extension to ScPy.
-
-As Python is an easily extendible general purpose language, it has applications in many scientific
-fields. Glover et al.~\cite{glover2011python} present a discussion of the use of Python for audio
-signal processing. They note that while audio processing libraries exist for many languages,
-scripting languages such as Python or MATLAB see widespread adoption because of the speed at which
-prototypes can be built. General purpose Python libraries such as NumPy, SciPy and Matplotlib
-include many essential functions for audio processing, such as file IO, Fourier transforms, signal
-processing and interpolation. There are additionally more specific libraries such as Simpl, a
-sinusoidal modelling library, and Modal, an onset detection library. Any of these libraries can
-easily be used within ScPy simply by installing and importing them.
-
-Another open-source Python library which could be used in ScPy is librosa~\cite{mcfee2015librosa}
-by McFee et al.\ librosa is a music signal analysis library designed to help transition music
-information retrieval (MIR) researchers into using Python, as well as make MIR techniques readily
-available to all Python programmers.\ librosa's core design choices and conventions are documented
-in detail to inspire future library authors to follow similarly good coding practice. As well as
-standard audio analysis functions, librosa also includes support for effects and audio IO.\@
-
-librosa has quite an extensive list of features and would likely cover all of the audio related
-functionality which is missing from NumPy. It can be used with ScPy simply by installing and importing
-it.
-
-The ixi lang~\cite{magnusson2011ixi} is a live coding language. It is implemented as an interpreter
-built in SuperCollider, a choice which allows for SuperCollider and ixi lang code to be mixed
-together in a similar manner to ScPy. The ixi lang is built to reduce the startup time in a live
-coding performance and enable faster musical experimentation through syntax more concise than
-SuperCollider. The paper also contains user feedback on the usability of the ixi lang.
-
-Since the ixi lang is similar to ScPy in many ways, many of the design choices made in ixi lang are
-also appropriate in ScPy. Additionally, lessons learned from the user feedback for ixi lang could
-potentially be used to improve usability of ScPy.
 
@@
 extension could add the possibility to experiment with many novel effects.
+
+Klingbeil~\cite{klingbeil2005software} describes the techniques used in a high quality graphical
+spectral editing and resynthesis program called SPEAR\@. The FFT analysis is done taking into
+account the fundamental frequency of the sound to analyze, so that the transform can be optimized
+to accurately capture as many partials as possible. Spectral peaks are then selected and very quiet
+peaks are discarded. A linear prediction method is used to follow the peaks over time. The data
+structure for storage of spectral data involves division of time into a number of frames which keep
+track of which partials are active during that time, so that the search for a specific partial can
+be narrowed. User interface design choices are also covered in detail.
+
+Although the interface and granularity of SPEAR differs significantly from those in ScPy, the
+techniques for high quality resynthesis presented should prove useful for improving the quality of
+the sound output.
+
+While the Fourier transform is by far the most commonly used technique for spectral manipulation in
+digital audio effects, there exist numerous other transformations which may be useful as well to
+achieve certain audio effects. One such technique is the Mellin transform, described in a audio
+effects context by De Sena and Rocchesso~\cite{de2004fast}. The Mellin transform has a scale
+invariance property that is analogous to the shift invariance property of the Fourier transform. It
+can be computed quickly by applying two extra steps before doing a Fourier transform: exponential
+time warping and exponential multiplication. A few simple audio effects are presented using the
+Mellin transform, time stretching, filtering and random phase deviation (whisperization).
+
+As the Mellin transform can be implemented simply as an extension to the Fourier transform, it
+should be fairly easy to add support for it in ScPy, and it will hopefully enable the use of
+some different audio effects than can be achieved easily with the Fourier transform. However, the
+paper mentions that a short-time Mellin transform has not been attempted, so some further work may
+need to be done in that area first.
+
+Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
+code to generate sound is written live during a performance, often on a laptop computer. Although
+the use graphical programming environments such as Reaktor, Pure Data, or MAX/MSP in live
+performances is in some sense live coding, the paper focuses on the use of text based programming
+environments. Many of the pros and cons of live coding are presented. Although live coding adds
+some very interesting new possibilities to live music performance, it is not without it's
+challenges. Case studies of two live coding environments are included: Slub and SuperCollider with
+the just in time library.
+
+ScPy can be used for live coding of audio effects, similarly to how the environments studied in
+this paper can be used to live code synths and pattens. More details about how ScPy was designed to
+facilitate this can be found in \autoref{sec:methodology}.
+
+Live coding presents an interesting contrast to contemporary music with many clear differences.
+Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
+practice. While some argue that live coding must be performed live in front of an
+audience, an alternate perspective suggests that time constraints are not necessary and that the
+practice can refer to any improvisational music created by programming. The practice of live coding
+is still quite young --- it is mentioned that professional violinists will typically have practiced
+for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
+hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
+extensive programming experience. Another thing mentioned is the high level of abstraction offered
+by live coding creates a large startup time for producing meaningful output from a blank slate.
+Additionally, many exercises are presented for improving one's live coding ability.
+
+The ideas in Nilson's paper shed light on some of the challenges faced in this project.
+One such challenge is the amount of writing required to produce an interesting result from a blank
+slate. We intend to minimise this by providing many effect primitives and composition
+mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
+quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
+
+Domain-Specific Languages (DSLs) are small programming languages designed for use in a specific
+problem domain. They often lack general purpose programming constructs and have a syntax which is
+very concise for describing programs in their domain. Van Deursen et al.~\cite{van2000domain}
+present a survey of available DSL literature, along with the pros and cons of using a DSL, example
+DSLs, design methodology and implementation approaches. The biggest advantage of DSLs is that they
+enable expression of solutions in a way that is common for the problem domain. However, they can be
+costly to develop and difficult to design. Implementation techniques include the classical approach
+of interpretation or compilation, embedded languages / domain-specific libraries, preprocessing, or
+compiler extension. Additionally, the references of the paper each include a short summary.
+
+This review reveals many of the possible options for the DSL used within the project. Although we
+decided not to design a DSL for our project, analyzing the design and implementation techniques
+presented here will be helpful for future work in this area.
+
+As Python is an easily extendible general purpose language, it has applications in many scientific
+fields. Glover et al.~\cite{glover2011python} present a discussion of the use of Python for audio
+signal processing. They note that while audio processing libraries exist for many languages,
+scripting languages such as Python or MATLAB see widespread adoption because of the speed at which
+prototypes can be built. General purpose Python libraries such as NumPy, SciPy and Matplotlib
+include many essential functions for audio processing, such as file IO, Fourier transforms, signal
+processing and interpolation. There are additionally more specific libraries such as Simpl, a
+sinusoidal modelling library, and Modal, an onset detection library. Any of these libraries can
+easily be used within ScPy simply by installing and importing them.
+
+Another open-source Python library which could be used in ScPy is librosa~\cite{mcfee2015librosa}
+by McFee et al.\ librosa is a music signal analysis library designed to help transition music
+information retrieval (MIR) researchers into using Python, as well as make MIR techniques readily
+available to all Python programmers.\ librosa's core design choices and conventions are documented
+in detail to inspire future library authors to follow similarly good coding practice. As well as
+standard audio analysis functions, librosa also includes support for effects and audio IO.\@
+
+librosa has quite an extensive list of features and would likely cover all of the audio related
+functionality which is missing from NumPy. It can be used with ScPy simply by installing and importing
+it.
+
+The ixi lang~\cite{magnusson2011ixi} is a live coding language. It is implemented as an interpreter
+built in SuperCollider, a choice which allows for SuperCollider and ixi lang code to be mixed
+together in a similar manner to ScPy. The ixi lang is built to reduce the startup time in a live
+coding performance and enable faster musical experimentation through syntax more concise than
+SuperCollider. The paper also contains user feedback on the usability of the ixi lang.
+
+Since the ixi lang is similar to ScPy in many ways, many of the design choices made in ixi lang are
+also appropriate in ScPy. Additionally, lessons learned from the user feedback for ixi lang could
+potentially be used to improve usability of ScPy.
 

### 2016-08-28 doc/paper.tex
extras

@@
 
-ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest code
-and documentation is available on GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest
+code, documentation, and issues are available on
+GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}. Contributions to the project are
+welcome.
 
@@
 
+The choice of SuperCollider was largely because it is one of the most widely used audio programming
+environments. Using SuperCollider also enabled the reuse of many parts of its codebase. Since it
+already has STFT based operations, it was a good target for extension purposes in that area.
+
 Although designing a domain-specific language for this task was briefly considered, we eventually
@@
 
+Complex numbers are used extensively, even to represent non-Cartesian data such as polar
+coordinates or phase vocoder bins in ScPy. In polar form, the real and imaginary parts represent
+magnitude and phase, respectively. Similarly for the phase vocoder, real and imaginary parts
+represent magnitude and frequency. This is done to take advantage of the limited data types
+available in NumPy and relative convenience of working with complex numbers as opposed to tuples or
+classes. Future work could extend NumPy to give more descriptive names to these data fields.
+
 \begin{listing}[ht]
@@
     x = pv.to_bin_offset(x)
-    x.imag = -x.imag
+    x.imag = -x.imag # imaginary part holds frequency
     x = pv.from_bin_offset(x)
@@
 information must be encoded as well. However, just passing the type of a variable is not enough to
-know what we can do with it, since we also want to know if it inherits from a class we can use.
+know what we can do with it, since we also want to know if it inherits from a class we can use. The
+encoding and decoding implementation is unfortunately fairly complex, but creating a UGen with
+inputs this complicated was likely not considered during the design of SuperCollider, so some
+friction was to be expected in this area.
 
@@
 
+Throughout~\autoref{sec:lit}, articles are discussed which contain processing techniques that could
+be used in ScPy. Although implementations of many of these techniques are likely to become a part
+of ScPy in the future, there is no obstacle to the reader implementing them as Python libraries
+usable by ScPy or even as open-source contributions to the project.
+
 \newpage\printbibliography{}

### 2016-08-29 doc/paper.tex
split lit review

@@
 \addbibresource{references.bib}
-
-\graphicspath{{images/}}
 
@@
 
+\subsection{NIME}
+
 With any New Interface for Musical Expression (NIME), it is important to evaluate how it performs
@@
 
+\subsection{Spectral audio processing}
+
 The phase vocoder is an essential part of spectral audio manipulation. De G{\"o}tzen et
@@
 the implementation in the future.
-
+\\\par
 Keiler and Sylvain~\cite{keiler2002survey} present a survey of 6 techniques for accurate analysis
@@
 not clear if the trade-off will be worthwhile.
-
+\\\par
 Audio morphing is defined as a gradual change from one sound to the other. While morphing can be
@@
 techniques could be implemented in Python as a future extension to ScPy.
-
+\\\par
 Laroche and Dolson~\cite{laroche1999improved} present an improvement to the phase vocoder which
@@
 could implement many of the proposed modifications to the phase vocoder to improve sound quality.
-
+\\\par
 Bradford et al.\ have published a number of works detailing the sliding
@@
 extension could add the possibility to experiment with many novel effects.
-
+\\\par
 Klingbeil~\cite{klingbeil2005software} describes the techniques used in a high quality graphical
@@
 the sound output.
-
+\\\par
 While the Fourier transform is by far the most commonly used technique for spectral manipulation in
@@
 need to be done in that area first.
-
+\\\par
 Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
@@
 facilitate this can be found in \autoref{sec:methodology}.
+
+\subsection{Lanugages}
 
@@
 quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
-
+\\\par
 Domain-Specific Languages (DSLs) are small programming languages designed for use in a specific
@@
 presented here will be helpful for future work in this area.
-
+\\\par
 As Python is an easily extendible general purpose language, it has applications in many scientific
@@
 easily be used within ScPy simply by installing and importing them.
-
+\\\par
 Another open-source Python library which could be used in ScPy is librosa~\cite{mcfee2015librosa}
@@
 it.
-
+\\\par
 The ixi lang~\cite{magnusson2011ixi} is a live coding language. It is implemented as an interpreter
@@
 The choice of SuperCollider was largely because it is one of the most widely used audio programming
-environments. Using SuperCollider also enabled the reuse of many parts of its codebase. Since it
+environments. Using SuperCollider also enabled the reuse of many parts of its code base. Since it
 already has STFT based operations, it was a good target for extension purposes in that area.

### 2016-08-29 doc/paper.tex
shuffling, editing, conclusion

@@
 within SuperCollider, it is possible to do numerical processing operations which would be too slow
-to perform in real time with SuperCollider alone. Through access to the NumPy library, previously
+to perform in real-time with SuperCollider alone. Through access to the NumPy library, previously
 difficult or impossible phase vocoder operations become trivial. Although our usage of this library
@@
 ScPy is developed and maintained as an open-source project under the GPLv3 license. The latest
-code, documentation, and issues are available on
-GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}. Contributions to the project are
-welcome.
+code, documentation, examples, and issues are available on
+GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}. The code has been developed and
+tested on the Linux operating system using SuperCollider version 3.6.6.
 
@@
 developed by Birnbaum et al.\ can be seen in~\autoref{fig:nime}.
-
 \begin{figure}[ht]
@@
 \end{figure}
+\\\par
+Live coding presents an interesting contrast to contemporary music with many clear differences.
+Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
+practice. While some argue that live coding must be performed live in front of an
+audience, an alternate perspective suggests that time constraints are not necessary and that the
+practice can refer to any improvisational music created by programming. The practice of live coding
+is still quite young --- it is mentioned that professional violinists will typically have practiced
+for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
+hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
+extensive programming experience. Also mentioned is the high level of abstraction offered
+by live coding creates a large startup time for producing meaningful output from a blank slate.
+Additionally, many exercises are presented for improving one's live coding ability.
+
+The ideas in Nilson's paper shed light on some of the challenges faced in this project.
+One such challenge is the amount of writing required to produce an interesting result from a blank
+slate. We intend to minimise this by providing many effect primitives and composition
+mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
+quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
 
@@
 These new algorithms are efficient implementations for the STFT with a hop size of a single sample.
-Although such a small hop size would be impossible to process in real time using current hardware
+Although such a small hop size would be impossible to process in real-time using current hardware
 and the STFT, the sliding DFT makes it possible using consumer GPU
@@
 Using Theano, the sliding DFT and sliding phase vocoder could be implemented easily in ScPy. It
-remains unknown if it will be possible to run them in real time, but even if not this future
+remains unknown if it will be possible to run them in real-time, but even if not this future
 extension could add the possibility to experiment with many novel effects.
@@
 need to be done in that area first.
-\\\par
+
+\subsection{Lanugages}
+
 Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
@@
 facilitate this can be found in \autoref{sec:methodology}.
-
-\subsection{Lanugages}
-
-Live coding presents an interesting contrast to contemporary music with many clear differences.
-Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
-practice. While some argue that live coding must be performed live in front of an
-audience, an alternate perspective suggests that time constraints are not necessary and that the
-practice can refer to any improvisational music created by programming. The practice of live coding
-is still quite young --- it is mentioned that professional violinists will typically have practiced
-for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
-hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
-extensive programming experience. Another thing mentioned is the high level of abstraction offered
-by live coding creates a large startup time for producing meaningful output from a blank slate.
-Additionally, many exercises are presented for improving one's live coding ability.
-
-The ideas in Nilson's paper shed light on some of the challenges faced in this project.
-One such challenge is the amount of writing required to produce an interesting result from a blank
-slate. We intend to minimise this by providing many effect primitives and composition
-mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
-quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
 \\\par
@@
 
-Use of ScPy is through the classes \texttt{Py} and \texttt{PyOnce}. \texttt{Py} is used for
-processing real-time data, and \texttt{PyOnce} is used for doing initialization or other one time
-processing. Both have one required argument, a block of Python code. A map of variable names may
-optionally be provided to bind SuperCollider variables to Python variables. Additionally, there is
-an optional \texttt{DoneAction} argument which specifies an action to occur in SuperCollider after
-the Python code has finished executing.
+Use of ScPy is through the classes \texttt{Py}, \texttt{PyOnce}, \texttt{PyFile}, and
+\texttt{PyOnceFile}. \texttt{Py} is used for processing real-time data, and \texttt{PyOnce} is used
+for doing initialization or other one time processing. Both have one required argument, a block of
+Python code. The \texttt{PyFile} and \texttt{PyOnceFile} variants take a filename instead of inline
+code. For all classes, a map of variable names may optionally be provided to bind SuperCollider
+variables to Python variables. Additionally, there is an optional \texttt{DoneAction} argument
+which specifies an action to occur in SuperCollider after the Python code has finished executing.
 
@@
 Since UGens may only be used as part of a synth, \texttt{PyOnce} is provided as a wrapper which
-hides this detail from the user and allows the execution of Python code outside of a synth. It is
-useful for doing initialization work. Since the current implementation executes Python code in a
-single global namespace, a typical usage patten is for state to be defined in a \texttt{PyOnce}
-block which is later accessed and modified in a \texttt{Py} block.
+hides this detail from the user and allows the execution of Python code outside of a synth. It
+internally just starts a synth with the \texttt{doneAction} argument set such that the synth
+terminates after running once.
 
@@
 of ScPy in the future, there is no obstacle to the reader implementing them as Python libraries
-usable by ScPy or even as open-source contributions to the project.
+usable by ScPy or even as contributions to the project.
 
-\newpage\printbibliography{}
+\section{Conclusion}
+
+While the number of mentioned future extensions to ScPy is exhausting to consider, the basic
+implementation that currently exists is more than enough to yield countless exotic and novel audio
+effects. In the limited experiments that have been done using the software, it was found to be very
+easy to produce many advanced sounds, thanks to the excellent documentation and implementation of
+the Python language and NumPy library. Although it was certainly possible to make these effects
+before, ScPy makes it possible to experiment with them at an unprecedented rate and listen to their
+output in real-time.
+
+Although initial tests have been promising, they have so far have had a focus on spectral
+processing, leaving time domain audio and control rate data processing largely untested. Further
+experiments in these areas could reveal previously unknown new use cases for the project.
+
+However, the code may contain bugs due to the infancy of the project and the nonstandard ways we are
+interfacing with SuperCollider. Users are encouraged to submit bug reports on the project's GitHub
+page\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+
+\printbibliography{}
 

### 2016-08-29 doc/paper.tex
little fixes

@@
 
-\author{Noah Weninger \and Abram Hindle}
+\author{Noah Weninger\footnote{\href{mailto:nweninge@ualberta.ca}{nweninge@ualberta.ca}}\hskip
+0.75em and Abram
+Hindle\footnote{\href{mailto:hindle1@ualberta.ca}{hindle1@ualberta.ca}}\\Department of Computing
+Science\\University of Alberta\\Edmonton, Canada}
 
@@
 
-However, the code may contain bugs due to the infancy of the project and the nonstandard ways we are
-interfacing with SuperCollider. Users are encouraged to submit bug reports on the project's GitHub
-page\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+Please note that the code may contain bugs due to the infancy of the project and the nonstandard
+ways we are interfacing with SuperCollider. Users are encouraged to submit bug reports on the
+project's GitHub page\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
 

### 2016-08-29 doc/paper.tex
proofreading & editing run 1

@@
 \usepackage{tkz-kiviat}
+\usepackage[margin=1in]{geometry}
 
@@
 
-Although a number of open-source projects which interface other languages with SuperCollider exist
-\cite{systemsinterfacingwithsc,magnusson2011ixi,orlarey2009faust}, they are mostly SuperCollider
-clients, which have features to control synths and send other automation messages. One notable
-exception is OctaveSC~\cite{octavesc}, which embeds GNU Octave within SuperCollider. However, it
-is designed for performing operations on control rate data matrices and suffers from performance
-issues which prevent it from being used on audio rate data.
+Although a number of open-source projects which interface other languages with SuperCollider (SC)
+exist~\cite{systemsinterfacingwithsc,magnusson2011ixi,orlarey2009faust}, they are mostly
+SuperCollider clients, which have features to control synths and send other automation messages.
+One notable exception is OctaveSC~\cite{octavesc} which embeds GNU Octave within SuperCollider.
+However, it is designed for performing operations on control rate data matrices and suffers from
+performance issues which prevent it from being used on audio rate data.
 
 One of the goals of this project is to enable more flexible experimentation with FFT and phase
-vocoder based operations. SuperCollider contains a total of 32 built in spectral operations,
+vocoder based operations. SuperCollider contains a total of 32 built in spectral operations
 which provide the ability to produce many simple effects. There are additionally some extra
@@
 GitHub\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}. The code has been developed and
-tested on the Linux operating system using SuperCollider version 3.6.6.
+tested on the Linux operating system using SuperCollider version 3.6.6 and Python version 3.5.2.
 
@@
 
-With any New Interface for Musical Expression (NIME), it is important to evaluate how it performs
+With any New Interface for Musical Expression (NIME) it is important to evaluate how it performs
 and is interacted with across various dimensions. Birnbaum et al.~\cite{birnbaum2005towards}
@@
 installations tend to occupy the left. Although this model performs well on the instruments
-evaluated in the paper, it is based on subjective evaluation and it is likely that it does not
-adequately cover the space of all possible instruments.
+evaluated in the paper, it is based on subjective evaluation and may accurately represent the entire
+space of possible NIMEs.
 
-ScPy is unlike many NIMEs because it does not have the properties of any traditional instrument.
-However, it can still be considered a NIME because it ultimately is a tool for musical expression,
+ScPy is unlike many NIMEs because it is not used like any traditional musical instrument.
+However, it can still be considered a NIME because it is a tool to enhance musical expression,
 just with an unconventional method of interaction. A plot of ScPy in the 7-dimensional space
 developed by Birnbaum et al.\ can be seen in~\autoref{fig:nime}.
+
 \begin{figure}[ht]
@@
 \end{figure}
-\\\par
-Live coding presents an interesting contrast to contemporary music with many clear differences.
-Nilson~\cite{nilson2007live} presents a review of the challenges of live coding
-practice. While some argue that live coding must be performed live in front of an
-audience, an alternate perspective suggests that time constraints are not necessary and that the
-practice can refer to any improvisational music created by programming. The practice of live coding
-is still quite young --- it is mentioned that professional violinists will typically have practiced
-for 10000 hours, where as an email survey of live coders revealed most had dedicated about 100
-hours. Of course, the prerequisite knowledge for live coding may be higher; many live coders have
-extensive programming experience. Also mentioned is the high level of abstraction offered
-by live coding creates a large startup time for producing meaningful output from a blank slate.
-Additionally, many exercises are presented for improving one's live coding ability.
 
-The ideas in Nilson's paper shed light on some of the challenges faced in this project.
-One such challenge is the amount of writing required to produce an interesting result from a blank
+Nilson~\cite{nilson2007live} presents a review of the challenges of live coding practice. Live
+coding presents an interesting contrast to contemporary music with many clear differences.  While
+some argue that live coding must be performed live in front of an audience, an alternate
+perspective suggests that time constraints are not necessary and that the practice can refer to any
+improvisational music created by programming. The practice of live coding is still quite new.
+For example, Nilson writes that professional violinists will typically have practiced for 10000 hours, where
+as an email survey of live coders revealed most had dedicated about 100 hours. Of course, the
+prerequisite knowledge for live coding may be higher; many live coders have extensive programming
+experience. Also mentioned is the high level of abstraction offered by live coding creates a large
+startup time for producing meaningful output from a blank slate.  Additionally, many exercises are
+presented for improving one's live coding ability.
+
+Nilson's work sheds light on some of the challenges faced in this project.
+One such challenge is the amount of code required to produce an interesting result from a blank
 slate. We intend to minimise this by providing many effect primitives and composition
-mechanisms. However, it will be important to ensure the language is accessible to ensure it can be
+mechanisms. However, it will be important to ensure the language is accessible so it can be
 quickly learned to a level of expertise where lack of technical knowledge is not an impediment.
@@
 tricks to make the implementation faster or simpler. The implementation is split into two
-functions, \texttt{pv\_analyze} and \texttt{pv\_synthesize}. These functions operate as a tightly
+functions: \texttt{pv\_analyze} and \texttt{pv\_synthesize}. These functions operate as a tightly
 coupled set due to the phase information maintained between frames. One of the techniques presented
@@
 the implementation in the future.
-\\\par
+
 Keiler and Sylvain~\cite{keiler2002survey} present a survey of 6 techniques for accurate analysis
 of stationary sounds. The first method, plain FFT, uses a Hann window and assumes the frequencies
-of the sinusoids at centered within each bin. Parabolic interpolation notices that the shape of a
-magnitude peak looks like a parabola, and thus attempts to fit a parabola to the magnitudes of bins
+of the sinusoids are centered within each bin. Parabolic interpolation notices that the shape of a
+magnitude peak looks like a parabola and thus attempts to fit a parabola to the magnitudes of bins
 nearby each local maxima, yielding more accurate results than a plain FFT\@. The triangle algorithm
@@
 not clear if the trade-off will be worthwhile.
-\\\par
+
 Audio morphing is defined as a gradual change from one sound to the other. While morphing can be
-achieved to some extent simply by cross fading two signals, perceptually this sounds simply like
+achieved to some extent simply by cross fading two signals; perceptually this sounds simply like
 one sound is becoming quieter as the other is becoming louder. Sethares and Bucklew
@@
 convincing morph operation where there is no clear separation between the two sounds at any point
-during the morphing process. The method allows the use of different kernel functions to provide
+during the morphing process. This method allows the use of different kernel functions to provide
 different kinds of morphing. These include \textit{Cross-fade}, \textit{Heat equation}, and
@@
 techniques could be implemented in Python as a future extension to ScPy.
-\\\par
+
 Laroche and Dolson~\cite{laroche1999improved} present an improvement to the phase vocoder which
-aims to prevent the ``phasiness'' problem where different parts of a sound lose their phase
+aims to prevent the \textit{phasiness} problem where different parts of a sound lose their phase
 coherence, as well as provide a significant decrease in computational cost. The paper recognizes
@@
 ensure that the reproduced sound is without artifacts: horizontal and vertical. The basic phase
-vocoder algorithm is built to preserve horizontal phase coherence --- but it does so at the expense
+vocoder algorithm is built to preserve horizontal phase coherence but it does so at the expense
 of vertical phase coherence. To remedy this, they build on prior work by presenting two new
@@
 exotic audio effects. The core of the techniques is a peak detection process followed by peak
-shifting. Since these new techniques operate on peaks, the standard phase vocoder algorithm can be
+shifting. Since these new techniques operate on peaks the standard phase vocoder algorithm can be
 simplified to not require phase unwrapping, reducing computational complexity. Additionally, they
@@
 could implement many of the proposed modifications to the phase vocoder to improve sound quality.
-\\\par
+
 Bradford et al.\ have published a number of works detailing the sliding
@@
 hardware~\cite{bradford2011real}. The sliding phase vocoder additionally has a number of advantges
-over traditional implementations. Quality is greatly improved (at a high computational cost). Each
+over traditional implementations. Quality is greatly improved but at a high computational cost. Each
 analysis bin can hold any frequency, essentially making the spectrum more of a generic oscillator
@@
 extension could add the possibility to experiment with many novel effects.
-\\\par
+
 Klingbeil~\cite{klingbeil2005software} describes the techniques used in a high quality graphical
 spectral editing and resynthesis program called SPEAR\@. The FFT analysis is done taking into
-account the fundamental frequency of the sound to analyze, so that the transform can be optimized
+account the fundamental frequency of the sound to analyze so that the transform can be optimized
 to accurately capture as many partials as possible. Spectral peaks are then selected and very quiet
@@
 structure for storage of spectral data involves division of time into a number of frames which keep
-track of which partials are active during that time, so that the search for a specific partial can
+track of which partials are active during that time so that the search for a specific partial can
 be narrowed. User interface design choices are also covered in detail.
@@
 the sound output.
-\\\par
+
 While the Fourier transform is by far the most commonly used technique for spectral manipulation in
@@
 time warping and exponential multiplication. A few simple audio effects are presented using the
-Mellin transform, time stretching, filtering and random phase deviation (whisperization).
+Mellin transform, time stretching, filtering and random phase deviation (\textit{whisperization}).
 
 As the Mellin transform can be implemented simply as an extension to the Fourier transform, it
-should be fairly easy to add support for it in ScPy, and it will hopefully enable the use of
+should be fairly easy to add support for it in ScPy and it will hopefully enable the use of
 some different audio effects than can be achieved easily with the Fourier transform. However, the
@@
 
-Collins et al.~\cite{collins2003live} present an introduction to the field of live coding, where
+Collins et al.~\cite{collins2003live} present an introduction to the field of live coding where
 code to generate sound is written live during a performance, often on a laptop computer. Although
@@
 challenges. Case studies of two live coding environments are included: Slub and SuperCollider with
-the just in time library.
+the \textit{just in time} library.
 
-ScPy can be used for live coding of audio effects, similarly to how the environments studied in
+ScPy can be used for live coding of audio effects similarly to how the environments studied in
 this paper can be used to live code synths and pattens. More details about how ScPy was designed to
 facilitate this can be found in \autoref{sec:methodology}.
-\\\par
+
 Domain-Specific Languages (DSLs) are small programming languages designed for use in a specific
@@
 of interpretation or compilation, embedded languages / domain-specific libraries, preprocessing, or
-compiler extension. Additionally, the references of the paper each include a short summary.
+compiler extension. The paper also includes an annotated bibliography.
 
@@
 presented here will be helpful for future work in this area.
-\\\par
+
 As Python is an easily extendible general purpose language, it has applications in many scientific
@@
 easily be used within ScPy simply by installing and importing them.
-\\\par
-Another open-source Python library which could be used in ScPy is librosa~\cite{mcfee2015librosa}
-by McFee et al.\ librosa is a music signal analysis library designed to help transition music
-information retrieval (MIR) researchers into using Python, as well as make MIR techniques readily
-available to all Python programmers.\ librosa's core design choices and conventions are documented
-in detail to inspire future library authors to follow similarly good coding practice. As well as
-standard audio analysis functions, librosa also includes support for effects and audio IO.\@
+
+Another audio related open-source Python library is librosa~\cite{mcfee2015librosa} by McFee et
+al.\ librosa is a music signal analysis library designed to help transition music information
+retrieval (MIR) researchers into using Python, as well as make MIR techniques readily available to
+all Python programmers.\ librosa's core design choices and conventions are documented in detail to
+inspire future library authors to follow similarly good coding practice. As well as standard audio
+analysis functions, librosa also includes support for effects and audio IO.\@
 
@@
 it.
-\\\par
-The ixi lang~\cite{magnusson2011ixi} is a live coding language. It is implemented as an interpreter
+
+The ixi lang~\cite{magnusson2011ixi} by Magnusson is a live coding language. It is implemented as an interpreter
 built in SuperCollider, a choice which allows for SuperCollider and ixi lang code to be mixed
@@
 coding performance and enable faster musical experimentation through syntax more concise than
-SuperCollider. The paper also contains user feedback on the usability of the ixi lang.
+SuperCollider. This paper also contains user feedback on the usability of the ixi lang.
 
 Since the ixi lang is similar to ScPy in many ways, many of the design choices made in ixi lang are
-also appropriate in ScPy. Additionally, lessons learned from the user feedback for ixi lang could
+also appropriate in ScPy. Lessons learned from the user feedback for ixi lang could
 potentially be used to improve usability of ScPy.
@@
 
-During the initial planning stage of the project, we wanted to essentially provide a set of
-fundamental operations on spectral data which could be composed together to perform any conceivable
-effect. It was also a requirement that these operations would be syntactically concise to enable
-easy and fast experimentation. Ideally, the system would be able to handle complex effects in
-real-time.
+During the initial planning stage of the project, we wanted to provide a set of fundamental
+operations on spectral data which could be composed together to perform any conceivable effect. It
+was also a requirement that these operations would be syntactically concise to enable easy and fast
+experimentation. Ideally, the system would be able to handle complex effects in real-time.
 
@@
 makes embedding it within other languages trivial in comparison to many other options. Second,
-Python's syntax is easily readable and concise --- it clearly adheres to our goals. Finally, Python
-has a massive number of community built libraries for performant scientific and numerical
+Python's syntax is easily readable and concise and therefore it adheres to our goals.  Finally,
+Python has a massive number of community built libraries for performant scientific and numerical
 computing. With a few small exceptions, which are clearly specific to spectral data (such as the
@@
 It may seem as though embedding an interpreted language within an interpreted language could offer
-no performance advantage. There is certainly some performance hit when switching languages and
+no performance advantage. There is certainly some loss of performance when switching languages and
 transferring data. However, the biggest advantage comes not strictly from performance but from the
 diversity of operations which become available with highly optimized implementations. NumPy is used
-internally by ScPy for handling data arrays, so the entire NumPy library of mathematical functions
+internally by ScPy for handling data arrays so the entire NumPy library of mathematical functions
 comes at no cost. Any other Python library can easily be imported as well. Limited experiments
@@
 Use of ScPy is through the classes \texttt{Py}, \texttt{PyOnce}, \texttt{PyFile}, and
-\texttt{PyOnceFile}. \texttt{Py} is used for processing real-time data, and \texttt{PyOnce} is used
-for doing initialization or other one time processing. Both have one required argument, a block of
+\texttt{PyOnceFile}. \texttt{Py} is used for processing real-time data and \texttt{PyOnce} is used
+for doing initialization or other one time processing. Both have one required argument: a block of
 Python code. The \texttt{PyFile} and \texttt{PyOnceFile} variants take a filename instead of inline
@@
     \begin{minted}[linenos=true]{SuperCollider}
-(PyOnce("
-    # comments in Python have different syntax!
-    print('Hello from Python!')
-    print('x is', x)
-", (x: 62)))
+PyOnce("print('Hello from Python!')");
     \end{minted}
@@
 ScPy is written mostly in C++, but with a small SuperCollider class library to connect the C++
-back-end, and a small Python library that provides some useful operations. Due to differences
+back-end and a small Python library that provides some useful operations. Due to differences
 between these languages, connecting them was somewhat awkward in some cases.
@@
 with a number of challenges. Essentially, all data must be serialized into an array of floats.
-Strings must be converted to an array of ASCII character codes, and prepended with their length.
-Arguments passed through to the Python code can have one of many types, and therefore type
+Strings must be converted to an array of ASCII character codes and prepended with their length.
+Arguments passed through to the Python code can have one of many types and therefore type
 information must be encoded as well. However, just passing the type of a variable is not enough to
@@
 inputs this complicated was likely not considered during the design of SuperCollider, so some
-friction was to be expected in this area.
+friction was expected in this area.
 
-Since \texttt{FFT} objects in SuperCollider operate on a single channel only, they must be placed
+Since \texttt{FFT} objects in SuperCollider operate on a single channel only they must be placed
 in \texttt{Array} objects to support multichannel audio. \texttt{Array}s can hold objects of any
@@
 \autoref{lst:boilerplate}. Although it would be possible to do this conversion
-automatically, it is left out for the purpose of not introducing a feature simply to work around an
+automatically it is left out for the purpose of not introducing a feature simply to work around an
 incomplete implementation of \texttt{FFT} objects in SuperCollider. The current conversions will
@@
 At the time of writing, ScPy only supports handling audio rate data via SuperCollider's
-\texttt{Buffer} objects. In order to make some use cases more ergonomic, ScPy could certainly be
+\texttt{Buffer} objects. In order to make some use cases more ergonomic ScPy could certainly be
 extended to support raw audio rate input. For example, this would make time domain audio data even
@@
 Multi-core processing support would allow for many more sounds to be processed in real-time.
-Currently, all processing is done on a single thread, even with many UGens running at once. Using
-the supernova server~\cite{blechmann2005supernova} may be helpful, however it has not yet been
+Currently, all processing is done on a single thread even with many UGens running at once. Using
+the supernova server~\cite{blechmann2005supernova} may be helpful however it has not yet been
 tested and ScPy may require some modifications to be usable with supernova.
@@
 Faust~\cite{orlarey2009faust} environments as well. It may be desirable to completely separate the
-embedded language from the host, in order to allow easier interoperation and a cleaner
+embedded language from the host in order to allow easier interoperation and a cleaner
 implementation.
@@
 
-While the number of mentioned future extensions to ScPy is exhausting to consider, the basic
-implementation that currently exists is more than enough to yield countless exotic and novel audio
-effects. In the limited experiments that have been done using the software, it was found to be very
-easy to produce many advanced sounds, thanks to the excellent documentation and implementation of
-the Python language and NumPy library. Although it was certainly possible to make these effects
-before, ScPy makes it possible to experiment with them at an unprecedented rate and listen to their
-output in real-time.
+While the number of future extensions to ScPy is numerous, the basic implementation that currently
+exists is more than enough to yield countless exotic and novel audio effects. In the limited
+experiments that we did using the software, it was found to be very easy to produce many advanced
+sounds, thanks to the excellent documentation and implementation of the Python language and NumPy
+library. Although it was possible to make these effects without our technology, ScPy makes it
+possible to experiment with them at an unprecedented rate and listen to their output in real-time.
 
-Although initial tests have been promising, they have so far have had a focus on spectral
-processing, leaving time domain audio and control rate data processing largely untested. Further
-experiments in these areas could reveal previously unknown new use cases for the project.
+Initial tests have been promising but they have so far have been focused on spectral processing,
+leaving time domain audio and control rate data processing largely untested. Further experiments in
+these areas could reveal previously unknown new use cases.
 
-Please note that the code may contain bugs due to the infancy of the project and the nonstandard
-ways we are interfacing with SuperCollider. Users are encouraged to submit bug reports on the
-project's GitHub page\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
+Those that wish to experiment with ScPy should note that the code may contain bugs due to the
+infancy of the project and the nonstandard ways we are interfacing with SuperCollider. It is our
+intention to maintain and improve this project in the future. Users are encouraged to submit bug
+reports on the project's GitHub page\footnote{\url{https://github.com/nwoeanhinnogaehr/ScPy}}.
 

### 2016-08-29 doc/paper.tex
edit run 2

@@
 prerequisite knowledge for live coding may be higher; many live coders have extensive programming
-experience. Also mentioned is the high level of abstraction offered by live coding creates a large
-startup time for producing meaningful output from a blank slate.  Additionally, many exercises are
-presented for improving one's live coding ability.
+experience. Also mentioned is that the high level of abstraction offered by live coding creates a large
+startup time for producing meaningful output from a blank slate. The paper presents many exercises
+for improving one's live coding ability.
 
@@
 
-Similar techniques~\cite{laroche1999new} enable new phase vocoder techniques that enable
-exotic audio effects. The core of the techniques is a peak detection process followed by peak
-shifting. Since these new techniques operate on peaks the standard phase vocoder algorithm can be
-simplified to not require phase unwrapping, reducing computational complexity. Additionally, they
-produce better quality output than the standard phase vocoder algorithm because of implicit
-phase-locking.
+Similar techniques also by Laroche and Dolson~\cite{laroche1999new} enable new phase vocoder
+techniques that can be used for exotic audio effects. The core of the techniques is a peak
+detection process followed by peak shifting. Since these new techniques operate on peaks the
+standard phase vocoder algorithm can be simplified to not require phase unwrapping, reducing
+computational complexity. Additionally, they produce better quality output than the standard phase
+vocoder algorithm because of implicit phase-locking.
 
@@
 multiplication in the frequency domain. Additionally, since the time resolution is so fine, effects
-that require time modulation such as FM or AM can now be performed in the frequency domain.
+that require high resolution modulation through time such as FM or AM can now be performed in the
+frequency domain.
 
@@
 time warping and exponential multiplication. A few simple audio effects are presented using the
-Mellin transform, time stretching, filtering and random phase deviation (\textit{whisperization}).
+Mellin transform: time stretching, filtering and random phase deviation (\textit{whisperization}).
 
@@
 environments. Many of the pros and cons of live coding are presented. Although live coding adds
-some very interesting new possibilities to live music performance, it is not without it's
+some very interesting new possibilities to live music performance, it is not without its
 challenges. Case studies of two live coding environments are included: Slub and SuperCollider with
@@
 
-This review reveals many of the possible options for the DSL used within the project. Although we
-decided not to design a DSL for our project, analyzing the design and implementation techniques
-presented here will be helpful for future work in this area.
+This review presents many useful ideas for language design within ScPy. Although we decided not to
+design a DSL for our project, analyzing the design and implementation techniques presented here
+will be helpful for future work in this area.
 
@@
 also appropriate in ScPy. Lessons learned from the user feedback for ixi lang could
-potentially be used to improve usability of ScPy.
+potentially be used to improve the usability of ScPy.
 
@@
     x = pv.to_bin_offset(x)
-    x.imag = -x.imag # imaginary part holds frequency
+    x.imag *= -1 # imaginary part holds frequency
     x = pv.from_bin_offset(x)
@@
 mutability. In practice, NumPy accepts a one element \texttt{ndarray} almost anywhere a
-\texttt{float} is accepted, so this detail is not an annoyance.
+\texttt{float} is accepted, so this detail is usually not an annoyance.
 
@@
 implemented in Python for performing frequency transformations. Interestingly, although all of
-SuperCollider's FFT operations are referred to as phase vocoder operations, there is actually no
-implementation of a phase vocoder to be found in the SuperCollider code base at the time of
-writing. It is likely that a phase vocoder was once intended to be implemented, but the 32
-so-called phase vocoder effects are simply transformations to the Cartesian or polar
-representations of the spectrum.
+SuperCollider's FFT operations are referred to as phase vocoder operations in the code and
+documentation, there is actually no implementation of a phase vocoder to be found in the
+SuperCollider code base at the time of writing. It is likely that a phase vocoder was once intended
+to be implemented, but the 32 so-called phase vocoder effects are simply transformations to the
+Cartesian or polar representations of the spectrum.
 
@@
 \texttt{Buffer} objects. In order to make some use cases more ergonomic ScPy could certainly be
-extended to support raw audio rate input. For example, this would make time domain audio data even
+extended to support raw audio rate input. This would make time domain audio data even
 simpler to work with than frequency domain data is currently. It would also enable SuperCollider's
@@
 Throughout~\autoref{sec:lit}, articles are discussed which contain processing techniques that could
-be used in ScPy. Although implementations of many of these techniques are likely to become a part
-of ScPy in the future, there is no obstacle to the reader implementing them as Python libraries
-usable by ScPy or even as contributions to the project.
+be used in ScPy. Although implementations of many of these techniques may to become a part of ScPy
+in the future, there is no obstacle to the reader implementing them as external Python libraries
+usable by ScPy.
 
@@
 
-While the number of future extensions to ScPy is numerous, the basic implementation that currently
-exists is more than enough to yield countless exotic and novel audio effects. In the limited
-experiments that we did using the software, it was found to be very easy to produce many advanced
-sounds, thanks to the excellent documentation and implementation of the Python language and NumPy
-library. Although it was possible to make these effects without our technology, ScPy makes it
-possible to experiment with them at an unprecedented rate and listen to their output in real-time.
+While the number of possible future extensions to ScPy is numerous, the basic implementation that
+currently exists is more than enough to yield countless exotic and novel audio effects. In the
+limited experiments that we did using the software, it was found to be very easy to produce many
+advanced sounds, thanks to the excellent documentation and implementation of the Python language
+and NumPy library. Although it was possible to make these effects without our technology, ScPy
+makes it possible to experiment with them at an unprecedented rate and listen to their output in
+real-time.
 
